### 2022-01-23 paper2.Rmd
Finish Results

---
title: "Article 2:	Information-Seeking Behavior Decision-Making between Online Search and Interpersonal Networks: The Case of Covid-19 Vaccinations "
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output:
  bookdown::pdf_document2: 
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "My Library.bib"
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      cache = TRUE,
                      fig.width = 7,
                      fig.asp = 0.8,
                      out.width = "80%",
                      fig.align="center"
                      # dev.args = list(png = list(type = "cairo"))
                      )

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, 
               # rms, 
               modelsummary, showtext, thematic, kableExtra, scales)

rain <- read_rds(here('data', 'RAiN_vacc_clean.rds')) %>% 
  mutate(across(where(is.logical), as.factor),
         date = lubridate::as_date(EndDate))

my_palette = c('#56b4e9','#999999','#E69F00','#009E73','#CC79A7','#0072B2','#D55E00','#E69F00')
my_palette  <- c(MetBrewer::met.brewer(name = 'Cross', type = 'discrete'), '#555F61')

## Automatically use showtext to render text
font_add_google("Lora", "lora")
showtext_auto()

theme_diss <- function(base_size = 14) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      # Figure assembly
      plot.title = element_text(family="lora",
                                size = rel(1), 
                                margin = margin(0,0,5,0), 
                                hjust = 0),
      plot.title.position = "plot",
      plot.subtitle = element_text(family="lora",
                                   size = rel(0.85)),
      plot.caption = element_text(family="lora",
                                  size = rel(0.70),
                                  hjust = 1),
      # Graphical Zones
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Axes
      axis.title = element_text(family="lora", 
                                size = rel(0.85)),
      axis.text = element_text(family="lora", 
                               size = rel(0.70)),
      axis.line = element_line(color = "black", 
                               arrow = arrow(length = unit(0.3, "lines"), 
                                             type = "closed")),
      # Legend
      legend.title = element_text(family="lora",
                                  size = rel(0.85)),
      legend.text = element_text(family="lora", 
                                 size = rel(0.70)),
      legend.key = element_rect(fill = "transparent", 
                                colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", 
                                       colour = NA),
      # Facetting 
      strip.background = element_rect(fill = "#17252D",
                                      color = "#17252D"),
      strip.text = element_text(family="lora",
                                size = rel(0.85), 
                                color = "white", 
                                margin = margin(5,0,5,0))
    )
}

theme_set(theme_diss())

conf_int_prop <- function(var){
  p <- mean(!!var, na.rm = T)
  q <- 1 - p
  n <- nrow(rain)
  min <- p - 1.96*sqrt((p*q)/n)
  max <- p + 1.96*sqrt((p*q)/n) 
  paste0(round(min, 2), ", ", round(max, 2))
}
print_perc <- function(var){
  label_percent()(mean(var, na.rm = T))
}

```

# Intro
It is unclear how individuals choose to search for information through the
multimedia environment of the 21st century. Much of the research conducted in
social networks and communication focus on senders, influencers, and persuasion
strategies [@mertonManifestLatentFunctions1968] rather than on “the receiver as
an active information seeker and processor."
[@johnsonComprehensiveModelCancerRelated1993].

Core discussion networks are the default lens to investigate network resources
and social support in surveys. The underlying assumption is that individuals
reach out to a handful of strong ties when in need of support, which can be
elicited in surveys using name generators [@marsdenCoreDiscussionNetworks1987].
This approach has yielded important insights, but largely overlooks crucial
processes of resource activation [@hurlbertCoreNetworksTie2000;
@perrySocialNetworkActivation2015; @smithDonPutMy2005]. Small
[-@smallSomeoneTalk2017] shows that the core discussion network does not capture
how people activate social support in practice and indicates that people draw on
much broader social connections for support, also calling back to the weak ties
research by Granovetter [-@granovetterStrengthWeakTies1973]. There is ongoing
research investigating how weak-tie support holds for different support types
based on the architecture put forward by House, Umberson, and Landis
[-@houseStructuresProcessesSocial1988] of instrumental, emotional, and
informational support. I follow the definition of Cohen and Wills
[-@cohenStressSocialSupport1985] and define informational support as “help in
defining, understanding,  and coping with problematic events and include
education, advice, or referral to another source of  support”
[@winemillerMeasurementStrategiesSocial1993, p. 640].

However, informational support can also be sought outside of the social network
context, namely via computer-mediated information search tools such as the
process of “Googling”. As the online environment began penetrating all facets of
modern human life, it makes sense that performing online search has become one
of the most convenient vehicles for information search. While Small
[-@smallSomeoneTalk2017] focuses on how support can depend on a network tie
simply "happened to be there", online search is theoretically the most
frictionless and costless mode of support. There are three computer-mediated
vehicles I consider for this analysis: search engines like Google, posting
questions on online forums like subreddits or Facebook groups, and posting a
status update online via Facebook or Twitter. The first two comprise what I call
“online search”, while the third is conceptually distinct and important to
include in the survey; yet, it is a hybrid of social network activation and
online search.

As online search or interaction also allows increased anonymity, no or lowered
social cost, and the potential avoidance of embarrassment and other negative
social interactions, I theorize that a majority of people who searched for
information will have done so online.

<!-- TODO add a hypothesis here -->

For this paper, I will use the case of Covid-19 vaccinations to explore the
variation among information search vehicles. There are theoretical reasons why
individuals would choose to search for information among their network ties
instead of online search. For instance, Rains
[-@rainsCopingIllnessDigitally2018] finds that patients tend to search for
technical information about an illness online but turn to their social network
for experiential information from others facing similar circumstances in the
case of cancer. Moreover, if an individual distrusts the medical establishment,
or popular media, they may be more likely to activate informational support
among their social network or turn to online groups that validate their
worldview [@bogersHowSocialAre2014].

This research project is important because the vehicles used to search by people
impacts the information they are exposed to. Given the large swaths of both
misinformation and disinformation regarding Covid-19 [@mottaHowRightLeaningMedia2020; @pathakInfodemicsCOVID19Role2020; @shahsavariConspiracyTimeCorona2020] and
the measurable impacts this misinformation has had on pandemic-related health
behaviors, how we choose to search for information affects our own lives and the
lives of others. Moreover, individuals who do not utilize social networks for
information search may feel like they have no network support, leading to
feelings of network isolation with major implications for mental health 
[@thoitsMechanismsLinkingSocial2011]

## section B: Research Questions
* How do computer-mediated or interpersonal network information-seeking strategies vary across populations? 
* What factors lead people to perform health information search online versus among social network ties? 
* How does the utilization of social networks, search engines, and social media as vary in the case of searching for information on the Covid-19 vaccine?

# Research Methodology
The data used for this research project are based on original survey data
sampled between `r format( min(rain$date), '%B %d, %Y')`  through 
`r format(max(rain$date), '%B %d, %Y')`. This survey was hosted on Qualtrics and
participants were paid and recruited using Amazon Turk (MTurk). The total valid
survey responses (n = `r nrow(rain)`) were selected from a total of 1,066
respondents; some responses were disqualified due to a few factors such as
detected usage of a VPN, random answer clicking (determined by illogical
responses), poor quality typed responses (e.g. social network alters
consistently named random nouns), or taking the survey more than once. The same
is slighted gender unbalanced, slightly skewing towards a male sample. For the
respondents who provided their zipcode (p = 
`r round(nrow(rain[!is.na(rain$region),]) / nrow(rain), 2)`), the 
sample is quite balanced with a slight skew towards MTurkers in the South (p = 
`r round(nrow(rain[rain$region == 'South',]) / nrow(rain), 2)`). 
The survey took an average of `r round(mean(rain$duration), 2)` minutes with a 
standard deviation of `r round(sd(rain$duration), 2)` minutes. The shortest 
valid survey took `r round(min(rain$duration), 2)` minutes and the longest took 
`r round(max(rain$duration), 2)` minutes. Participants were paid \$6.00 for their 
time after Amazon administrative fees, funded by a Grant awarded to Kelsey E.
Gonzalez and Nicolas Legewie by the Summer Institute in Computational Social
Science and the Russell Sage Foundation. The main portion of this survey
replicates the book, Someone to Talk To [@smallSomeoneTalk2017] by expandingon
Small’s finding that people draw on much broader sources than “important people”
for support by segregating weak-tie social support into instrumental, emotional,
and informational support (see House et al. 1988)). The latter part of the
survey focuses on resource activation for the specific case-study of information
during the Covid-19 pandemic and is analyzed in this paper.

```{r table-1-desc}
#| results = "asis"

desc_table <- rain %>% 
  select(receive_i_fr_dr,
  receive_i_fr_pers,
  receive_i_fr_tv,
  receive_i_fr_onlnet,
  receive_i_fr_onlgroup,
  int_sought,
  int_sought_fr_pers,
  int_sought_fr_dr,
  int_sought_fr_onlgroup,
  int_sought_fr_onlnet,
  int_sought_fr_onlsearch,
  int_sought_affect,
  vacc_status,
  vacc_view_pos,
  race_white,
  race_black,
  race_nat,
  race_asian, 
  hispanic,
  college,
  age, receive_i) %>% 
  as.data.frame()

datasummary(formula = 
  (`Received info` = receive_i) +
    (`Received info, Doctor` = receive_i_fr_dr) +
    (`Received info, Person` = receive_i_fr_pers) +
    (`Received info, News` = receive_i_fr_tv) +
    (`Received info, Social Networking Site` = receive_i_fr_onlnet) +
    (`Received info, Online Forum` = receive_i_fr_onlgroup) +
    (`Sought Info` = int_sought) +
    (`Sought Info, Doctor` = int_sought_fr_dr) +
    (`Sought Info, Person` = int_sought_fr_pers) +
    (`Sought Info, Social Networking Site` = int_sought_fr_onlnet) +
    (`Sought Info, Online Forum` = int_sought_fr_onlgroup) +
    (`Sought Info, Online Search` = int_sought_fr_onlsearch) +
    (`Vaccination Status` = vacc_status) +
    (`Vaccination Favorability` = vacc_view_pos) +
    (`Age` = age) +
    (`Hispanic or Latino/x` = hispanic) +
    (`Race, Black` = race_white) +
    (`Race, Black` = race_black) +
    (`Race, Native American` = race_nat) +
    (`Race, Asian or Pacific Islander` = race_asian) +
    (`Associate's Deg or above` = college)  ~ 
    (Mean + SD) ,
  data = desc_table,
  fmt = '%.2f',
  title = 'Descriptive Statistics for Dichotomous and Numeric Variables',
  notes =c(glue("Notes: {nrow(rain)} Surveyed, Conducted {format( min(rain$date), '%B %d')} through {format(max(rain$date), '%B %d')}, 2021.")),
  output = "kableExtra") %>% 
  kable_styling(latex_options = c("hold_position", "striped"))

```

## Measures

###  Sought Information
There are multiple dependent variables of interest for this paper. The first
variable is a dichotomous indicator of whether someone intentionally sought out
information about Covid-19 vaccinations. The survey asked, "And how about you 
yourself intentionally looking for information about a Covid-19 vaccine? Such 
information can include things such as advice, clarification, facts, and
experiences". About `r print_perc(rain$int_sought)` ($\widehat{p}$)
of respondents had intentionally sought out information about Covid-19 
vaccinations (95% Confidence Interval = \{`r conf_int_prop(rain$int_sought)`\}).

Further dependent variables segment the above question into multiple options;
"How did you look for information about the Covid-19 vaccine?". Respondents were
posed with 5 responses and 'other', of which they could select multiple. The
largest proportion, `r print_perc(rain$int_sought_fr_dr)`, asked their 
'doctor or another health professional' (95% Confidence Interval = 
\{`r conf_int_prop(rain$int_sought_fr_dr)`\}).
About `r print_perc(rain$int_sought_fr_pers)`  said they asked 'a person 
like friend, neighbor, or family member that [they] know' (95% Confidence 
Interval = \{`r conf_int_prop(rain$int_sought_fr_pers)`\}). An additional
`r print_perc(rain$int_sought_fr_onlsearch)` searched 'for [their] question 
using an online search engine such as Google or Bing' (95% Confidence Interval 
= \{`r conf_int_prop(rain$int_sought_fr_onlsearch)`\}). About a quarter of the 
sample, `r print_perc(rain$int_sought_fr_onlgroup)`, 'posted queries in an 
online discussion group, listserve, or other online forum like a Facebook Group 
or Subreddit' (95% Confidence Interval = \{`r conf_int_prop(rain$int_sought_fr_onlgroup)`\}).
And finally, `r print_perc(rain$int_sought_fr_onlnet)` of the sample 'posted queries 
on a social networking site such as Facebook timeline, Twitter status update, or 
Linkedin' (95% Confidence Interval = \{`r conf_int_prop(rain$int_sought_fr_onlnet)`\}).
Much of the question wording and the specific categories of search were inspired
by ICPSR project 37220 [@scanlon19].

```{r table-2-desc}
#| results = "asis"

datasummary(formula = (
  # (`Income` = income)+  
                         (`Gender` = gender)+  
                         (`Highest Education Level` = educ) + 
                         (`Plan to be Vaccinated if not` = vacc_status_future)  +
                         (`Found Info Sought Useful` = int_sought_useful) +
                         (`Census Region` = region)) ~ N + Percent(), 
 data = rain,
   fmt = '%.2f',
  title = 'Descriptive Statistics for Categorical Variables',
 notes =c(glue("Notes: {nrow(rain)} Surveyed, Conducted {format( min(rain$date), '%B %d')} through {format(max(rain$date), '%B %d')}, 2021.")),
 output = "kableExtra") %>%
  kable_styling(latex_options = c("hold_position", "striped"))

# TODO reorder factors in categorical table

```

### Vaccination Outlook

```{r}
vacc_status_freq <- rain[!is.na(rain$vacc_status_future),] %>% 
  group_by(col = vacc_status_future) %>% 
  summarise(n = n()) %>%
  mutate(n = n / sum(n))
```
Another Dependent variable for this analysis is a dichotomous variable
indicating whether the respondent has received a vaccination against Covid-19 or
plans to. For this survey, the number of doses were unimportant and actually
having already received the dose were irrelevant to the research question. This
variable instead focuses on intent or opinion about the vaccine. About 
`r print_perc(rain$vacc_view_pos)` have a positive view about the vaccine. This
construct is made through the combination of  two survey items. The first item
is actual vaccination status, obtained through the question 'Did you receive a
Covid-19 vaccine?'. `r print_perc(rain$vacc_status)` of the sample said that
they had indeed received a vaccination (95% Confidence Interval = \{
`r conf_int_prop(rain$vacc_status)`\}). As the national vaccination rate is
closer to 55% (CITE), this indicates some bias in our sample that must be
acknowledged: either MTurkers are more likely to be vaccinated that the normal
population (sampling bias) or there is some conformity bias in the responses
causing survey takers to provide false answers. The second survey item that
contributes to 'Vaccination Outlook' was only given to those who had not
responded 'yes' to the to their vaccination status. These respondents were asked 'Do you
plan to receive a vaccine for the prevention of the Covid-19 virus?'. Of those
who were asked this question (n = `r nrow(rain[!is.na(rain$vacc_status_future),])`),
`r label_percent()(filter(vacc_status_freq, col == 'Definitely yes') %>% pull(n))`
said they were 'definitely yes' getting the vaccine; 
`r label_percent()(filter(vacc_status_freq, col == 'Probably yes') %>% pull(n))` 
said they were 'probably yes' going to receive vaccine. The rest of the sample was more unsure: 
`r label_percent()(filter(vacc_status_freq, col == 'Might or might not') %>% pull(n))`
said they 'Might or might not', 
`r label_percent()(filter(vacc_status_freq, col == 'Probably not') %>% pull(n))`
said 'Probably not', and finally
`r label_percent()(filter(vacc_status_freq, col == 'Definitely not') %>% pull(n))`
said they would 'definitely not' receive the vaccine. The variable 'Vaccination
Outlook' indicates their either the respondent received a Covid-19 vaccine or
either probably or definitely will in the future.

### Independent Variables
The first independent variables used in this analysis are based on the question, 
'In the past 12 months, without searching for it, did you receive information
about the Covid-19 vaccine from … (check all sources you received information
from)'. This differs from the 'How did you look for information' question above
because it is based on passive reception of information. The largest proportion
of respondents had received information from a person like friend, neighbor, or
family member that [they] know ($\widehat{p}$ = `r print_perc(rain$receive_i_fr_pers)`
). Respondents also commonly received information from a television news channel
or a newspaper, about `r print_perc(rain$receive_i_fr_tv)` of the sample.
`r print_perc(rain$receive_i_fr_dr)` of the sample had received information from
their doctor or other health professional while `r print_perc(rain$receive_i_fr_onlnet)` 
received information from a social networking site such as Facebook timeline,
Twitter status update, or linkedin. Finally, the lowest proportion of the sample
had received information from an online discussion group, listserve, or other
online forum like a Facebook group or subreddit 
($\widehat{p}$ = `r print_perc(rain$receive_i_fr_onlgroup)`).

We also asked various demographic questions to understand our sample. The 
average age of our sample was `r round(mean(rain$age, na.rm = T), 2)`. 
Our sample was `r label_percent()(nrow(filter(rain, gender == "Female")) / nrow(rain))`
female, and `r label_percent()(nrow(filter(rain, gender == "Male")) / nrow(rain))`
male. The sample is diverse racially though some ethnic groups are under
represented compared to national averages. Race was evaluated using the
question, 'What is your race? If you are "mixed race", select all that apply.'
Because of the multi-selection question, each variable is dichotomous and
proportions are shown. `r print_perc(rain$race_white)` of the sample claimed
they were 'White', `r print_perc(rain$race_black)` claimed to be 'Black or
African American', `r print_perc(rain$race_nat)` chose 'American Indian or
Alaskan Native', `r print_perc(rain$race_asian)` selected 'Asian', and finally 
when asked ' Are you Hispanic, Latino/a/x, or Latin American Origin?', 
`r print_perc(rain$hispanic)` selected 'Yes'. Respondents were also asked to 
select the  highest level of education that you have completed. Based on the 
breakdown between 'Less than high school', 'High school graduate', 'Some college',
'Associate's or Technical degree', 'Bachelor's degree', and 'Graduate or 
professional degree',  `r print_perc(rain$college)` of the sample were classified
to have a College-level education (Associate's degree or above).

# Analysis

My analytic strategy proceeds in four steps. In Tables \@ref(tab:table-1-desc) 
& \@ref(tab:table-2-desc), I present descriptive statistics for all study 
variables, including variable ranges, means, and standard deviations. For this 
paper, I rely heavily on logistic regression, modeled in `r`; specifically, I 
fit a series of Binomial Generalized Linear Models [@venables2002a]. In Table \@ref(tab:table-model-1)^[All tables in this paper were created with the modelsummary package in `r` [@modelsummary] ]
, I fit a logistic regression model predicting
whether someone sought information about Covid-19 using their sources of
information ('Received info') as well as various demographic variables such as
age, education level and race. Figure \@ref(fig:plot-model-1) then illustrates
the relationship between the coefficients and information seeking.^[Figures in this paper were all created using the `r` packages jtools [@jtools] and ggplot2 [@wickham_etal; @wickham11] ].
In Table \@ref(tab:table-model-2), I disaggregate information seeking behaviors
by modelling a series of logistic regressions that predict the 5 different
avenues of seeking information about Covid-19: from another person, from a
doctor, on an online forum, on a social networking site, or using an online
search tool.
Finally, in Table \@ref(tab:table-model-3), I investigate whether there is a
relationship between information receiving and information seeking behaviors and
the choice to get vaccinated against Covid-19 through a logistic regression
model predicting Vaccination Outlook.  Figure \@ref(fig:plot-model-3) then illustrates
the relationship between the coefficients and vaccination outlook.

# Results
```{r clean_before_reg}
rain <- rain %>% 
  mutate(across(c(receive_i,
                  receive_i_fr_dr,
                  receive_i_fr_pers,
                  receive_i_fr_tv,
                  receive_i_fr_onlnet,
                  receive_i_fr_onlgroup,
                  int_sought,
                  int_sought_fr_pers,
                  int_sought_fr_dr,
                  int_sought_fr_onlgroup,
                  int_sought_fr_onlnet,
                  int_sought_fr_onlsearch,
                  int_sought_affect,
                  vacc_status,
                  vacc_view_pos,
                  hispanic,
                  race_white,
                  race_black,
                  race_nat,
                  race_asian, 
                  college), ~ factor(.x, levels = c(0, 1))))

regression_labels = c(
  'receive_i_fr_dr1' = 'Received Info, Doctor',
  'receive_i_fr_pers1' = 'Received Info, Person',
  'receive_i_fr_tv1' = 'Received Info, News',
  'receive_i_fr_onlnet1' = 'Received Info, Social Networking Site',
  'receive_i_fr_onlgroup1' = 'Received Info, Online Forum',
  'int_sought1' = 'Sought Info',
  'int_sought_fr_dr1' = 'Sought Info, Doctor',
  'int_sought_fr_pers1' = 'Sought Info, Person',
  'int_sought_fr_onlnet1' = 'Sought Info, Social Networking Site',
  'int_sought_fr_onlgroup1' = 'Sought Info, Online Forum',
  'int_sought_fr_onlsearch1' = 'Sought Info, Online Search',
  'age' = 'Age',
  'gender=Male' = 'Gender, Male',
  'gender=Other' = 'Gender, Other',
  'college1' = "Associate's Deg or above",
  'race_white1' = 'White',
  'race_black1' = 'Black',
  'race_nat1' = 'Native American',
  'race_asian1' = 'Asian',
  'hispanic1' = 'Hispanic or Latino/x')
```
## Information Seeking
```{r table-model-1}
#| results = "asis"

mod1 <- glm(formula = int_sought ~ 
              receive_i_fr_dr +   
              receive_i_fr_pers +   
              receive_i_fr_tv +   
              receive_i_fr_onlnet +   
              receive_i_fr_onlgroup +   
              age + 
              gender + 
              race_white +   
              race_black +    
              race_nat +      
              race_asian +  
              hispanic  + 
              college,
            data = rain, family = binomial(link="logit"))
nullmod1 <- glm(formula = int_sought ~ 1,
            data = rain, family = binomial(link="logit"))
psuedor2 <- as.numeric(1-(logLik(mod1)/logLik(nullmod1)))

rows <- tribble(~term, ~model1,
                "$R_{McFadden}^2$", psuedor2)
attr(rows, 'position') <- c(14, 4)

modelsummary(mod1,
            gof_omit = 'AIC', statistic = NULL,
            estimate = "{estimate}{stars} ({std.error})",
            title = 'What predicts someone intentionally searched for information',
            add_rows = rows,
            coef_map = regression_labels,
            notes = list('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Exponentiated Coefficients'),         
            exponentiate = TRUE,
            output = "kableExtra") %>%
  kable_styling(latex_options = c("hold_position", "striped"))

```

```{r plot-model-1}
#| fig.cap='Plot of Coefficients, Model 1'

plot_summs(mod1, 
           scale = TRUE,
           coefs =  c(
             'Received Info, Doctor' = 'receive_i_fr_dr',
             'Received Info, Person' = 'receive_i_fr_pers',
             'Received Info, News' = 'receive_i_fr_tv',
             'Received Info, Social Networking Site' = 'receive_i_fr_onlnet',
             'Received Info, Online Forum' = 'receive_i_fr_onlgroup',
             'Sought Info' = 'int_sought',
             'Sought Info, Doctor' = 'int_sought_fr_dr',
             'Sought Info, Person' = 'int_sought_fr_pers',
             'Sought Info, Social Networking Site' = 'int_sought_fr_onlnet',
             'Sought Info, Online Forum' = 'int_sought_fr_onlgroup',
             'Sought Info, Online Search' = 'int_sought_fr_onlsearch',
             'Age' = 'age',
             'Gender, Male' = 'gender=Male',
             'Gender, Other' = 'gender=Other',
             "Associate's Deg or above" = 'college',
             'White' = 'race_white',
             'Black' = 'race_black',
             'Native American' = 'race_nat',
             'Asian' = 'race_asian',
             'Hispanic or Latino/x' = 'hispanic'),
           robust = TRUE) +
  labs(title = 'What predicts someone intentionally searched for information',
       caption = 'Exponentiated Coefficients: Scaled and Robust') +
  ylab("") +
  theme_diss() +
  scale_color_manual(values = my_palette[3])
```

Table \@ref(tab:table-model-1) provides the results of my logistic regression
analysis predicting whether someone sought information about Covid-19. These
initial results suggest similar results for each vehicle of information
delivery: each mode of receiving information whether from a doctor, a person,
the news, a social networking site, or an online forum, all increase the odds of
seeking out information. For example, respondents who had received information 
from a doctor or other medical professional were `r round(exp(mod1$coefficients[['receive_i_fr_dr1']]),2)`
(relative odds ratio = $exp(\beta k)$, _p_ < 0.001) times more likely to seek
out information over someone who hadn't. Similarly, respondents who had received 
information from 'a person  like friend, neighbor, or family member that [they] know'
were `r round(exp(mod1$coefficients[['receive_i_fr_pers1']]),2)`
( _p_ < 0.001) times more likely to seek out information over someone who hadn't
received information from a person like that. Interesting, those who had
received information from an online discussion group, listserve, or other online
forum including a Facebook group or subreddit were 
`r round(exp(mod1$coefficients[['receive_i_fr_onlgroup1']]), 2)` ( _p_ < 0.001)
times more likely to search out more information than someone who hadn't. While 
most demographic variables don't seem to have a strong effect on whether someone
sought out information about the Covid-19 vaccine, a few key variables had strong
associations. Age, for instance, is associated with seeking information; every 
additional year of age a person holds they are
`r label_percent()(exp(mod1$coefficients[['age']])-1)` 
more likely to search for information, possibly indicating a generational effect
of information seeking patterns ( _p_ < 0.01). Additionally, those who were more 
educated, that is those with an associate's degree or higher, were 
`r round(exp(mod1$coefficients[['college1']]), 2)` 
more likely to seek additional information than those without the degree 
( _p_ < 0.05). And finally, there is evidence to suggest that the Black or 
African American sample was more likely to seek out information, specifically: 
Blacks were `r round(exp(mod1$coefficients[['race_black1']]), 2)` times more likely 
to seek out information over non-Blacks ( _p_ < 0.05). Figure \@ref(fig:plot-model-1) illustrates
the relationship between the coefficients and information seeking.

## Information Search Vehicles
```{r table-model-2}
#| results = "asis"

part2_models <- list()

part2_models[['Doctor']] <- glm(formula = int_sought_fr_dr ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Person']] <- glm(formula = int_sought_fr_pers ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Online Forum']] <- glm(formula = int_sought_fr_onlgroup ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Social Networking Site']] <- glm(formula = int_sought_fr_onlnet ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Online Search']] <- glm(formula = int_sought_fr_onlsearch ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

moda <- glm(formula = int_sought_fr_dr ~ 1,
            data = rain, family = binomial(link="logit"))
r2_a <- as.numeric(1-(logLik(part2_models[['Doctor']])/logLik(moda)))

modb <- glm(formula = int_sought_fr_pers ~ 1,
            data = rain, family = binomial(link="logit"))
r2_b <- as.numeric(1-(logLik(part2_models[['Person']])/logLik(modb)))

modc <- glm(formula = int_sought_fr_onlgroup ~ 1,
            data = rain, family = binomial(link="logit"))
r2_c <- as.numeric(1-(logLik(part2_models[['Online Forum']])/logLik(modc)))

modd <- glm(formula = int_sought_fr_onlnet ~ 1,
            data = rain, family = binomial(link="logit"))
r2_d <- as.numeric(1-(logLik(part2_models[['Social Networking Site']])/logLik(modd)))

mode <- glm(formula = int_sought_fr_onlsearch ~ 1,
            data = rain, family = binomial(link="logit"))
r2_e <- as.numeric(1-(logLik(part2_models[['Online Search']])/logLik(mode)))

rows <- tribble(~term, ~Doctor, ~Person, ~`Online Forum`, ~`Social Networking Site`, ~`Online Search`,
                "$R_{McFadden}^2$", r2_a,   r2_b, r2_c, r2_d, r2_e)
attr(rows, 'position') <- c(14, 9)

modelsummary(part2_models,
             align="llllll",
             statistic = NULL,
            estimate = "{estimate}{stars}",
            gof_omit = 'AIC',
            coef_map = regression_labels,
            notes = list('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Exponentiated Coefficients'),            
            title = 'What predicts someone intentionally searched for information, by vehicle of information search',
            exponentiate = TRUE,
            add_rows = rows,
            output = "kableExtra") %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "striped"))  %>%
  add_header_above(c(" " = 1, "Dependent Variable" = 5))
```

To further investigate how people search for information, I asked those who had
sought information out intentionally, "How did you look for information about
the Covid-19 vaccine?" Results are displayed in Table \@ref(tab:table-model-2).
Because respondents could select as many responses as applicable, each response
is recorded as a dichotomous variable and modeling using logistic regression.
Modeling each response as a separate formula is useful for interpretation, but
there are important caveats: because each logistic regression is on a different
scale, one must not compare the magnitude of coefficients between models. We
can, however, use the direction and significance of the coefficients to draw
conclusions about these different forms of information search vehicles.

The first model predicts whether or not a respondent sought information
intentionally about Covid-19 vaccinations by asking their doctor or another
health professional. First and foremost, I find that those who received
information about Covid-19 vaccinations were 
`r round(exp(part2_models$Doctor$coefficients[['receive_i_fr_dr1']]), 2)` times
more likely to seek out more information from their doctor than those who had 
not ( _p_ < 0.001). However, receiving information from other vehicles also 
increased the odds of seeking out information from medical professions. Those who
received information from a person like a friend or a relative were 
`r round(exp(part2_models$Doctor$coefficients[['receive_i_fr_pers1']]), 2)` times
more likely to ask their doctor ( _p_ < 0.001), while those who received 
information from the news or television were
`r round(exp(part2_models$Doctor$coefficients[['receive_i_fr_tv1']]), 2)`
times more likely ( _p_ < 0.01). I also find that those who received information
from an online forum like a subreddit were 
`r label_percent()(round(exp(part2_models$Doctor$coefficients[['receive_i_fr_onlgroup1']]), 2)-1)`
more likely to ask their doctors than those who hadn't ( _p_ < 0.05). Finally, 
respondents with an associate's degree or higher were 
`r label_percent()(round(exp(part2_models$Doctor$coefficients[['college1']]), 2)-1)`
more likely to ask their doctor for information on the Covid-19 vaccinations
than those who held less education ( _p_ < 0.01).

The second model predicts whether or not a respondent sought information about
Covid-19 vaccinations by asking a person like friend, neighbor, or family member
that they know. Only three coefficients in this model were significant and help
us draw any conclusions on who is most likely to activate their personal social
network for informational support. In this study, I find that receiving
information about the vaccine from a doctor, their personal network, or a social
networking site is associated with seeking out more information. Those who 
received information from a doctor or other medical professional were 
`r round(exp(part2_models$Person$coefficients[['receive_i_fr_dr1']]), 2)`
times more likely to seek information from a person like a friend  ( _p_ < 0.001). 
Moreover, it seems that receiving information from your personal social network,
like a friend or relative, is associated with seeking information from those 
same people; in fact, I find that those who did are 
`r round(exp(part2_models$Person$coefficients[['receive_i_fr_pers1']]), 2)` more
likely to do so ( _p_ < 0.001). Finally, the respondents who received information
about the vaccine on a social networking site were 
`r label_percent()(round(exp(part2_models$Person$coefficients[['receive_i_fr_onlnet1']]), 2)-1)`
 more likely to seek information from their personal network than those who hadn't 
( _p_ < 0.01). The unexpected result of this model is that very few factors predict
personal social network activation and this model has the lowest McFadden's R^2, 
indicating possible omitted variables.

I also wanted to see what was associated with posting in an online discussion
group, listserve, or other online forum like a Facebook Group or Subreddit about
Covid-19 vaccinations. This is conceptually different that posting a query on a
social networking site like Facebook or twitter where identities are involved,
but is theoretically more [costly?][friction?] than searching on an internet
search engine. I found that having received information about the vaccine from 
your personal social network (like a friend or relative), the news, or an online
forum was associated with higher odds of querying an online forum. Specifically,  
receiving information from your personal network is associated with 
`r round(exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_pers1']]), 2)`
higher odds ( _p_ < 0.001), from news or television media with `r round(exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_tv1']]), 2)` higher odds ( _p_ < 0.05), and 
receiving information on an online forum associated with 
`r round(exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_onlgroup1']]), 2)`
higher odds ( _p_ < 0.001). Moreover, I found that those with an associate's 
degree or above were 
`r round(exp(part2_models$'Online Forum'$coefficients[['college1']]), 2)` times
more likely to search for information on an online forum ( _p_ < 0.001). Finally,
I found a racial-ethnic with Hispanic-Latinos; I found that they were about 
`r label_percent()(exp(part2_models$'Online Forum'$coefficients[['hispanic1']])-1)`
more likely to query an online forum than non-Hispanics ( _p_ < 0.05). However, 
there were also a few variables that lowered the odds of posting in an online 
discussion group. First, those who had received information from a social 
networking site were
`r label_percent()(1-exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_onlnet1']]))` 
less likely to query a forum than those who hadn't (relative odds ratio =
$1 - exp(\beta k)$, _p_ < 0.01). Morever, I find more ethno-racial effects: 
Respondents who claimed Asian ancestry were 
`r label_percent()(1-exp(part2_models$'Online Forum'$coefficients[['race_asian1']]))` 
less likely to query online forum ( _p_ < 0.01).

The fourth model predicts whether an individual sought information about 
Covid-19 vaccinations through posted queries on a social networking site such 
as Facebook timeline, Twitter status update, or Linkedin. Receiving information
about Covid-19 seems to be quite related to posting on socia media. Those who 
received information from their doctor's had 
`r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_dr1']]), 2)`
higher odds of posting on a social networking site ( _p_ < 0.01), while those 
who received their information from tv/news had `r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_tv1']]), 2)` higher odds ( _p_ < 0.05). 
Furthermore, those who received information on social networking sites were more 
likely to post queries on social networking sites ($\beta$ = 
`r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_onlnet1']]), 2)`,  _p_ < 0.001). 
The respondents in our sample who received information on online forums had
`r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_onlgroup1']]), 2)` 
higher odds of posting queries on their social networking platforms ( _p_ < 0.001). 
I found a relationship between education and posting on social networking sites
as well; specifically, those with an Associate's degree or above were `r round(exp(part2_models$'Social Networking Site'$coefficients[['college1']]), 2)` times more likely to post vaccination 
queries on social networking sites than those with lower education ( _p_ < 0.001).
Finally, I find that Hispanic-Latinos were 
`r label_percent()(exp(part2_models$'Social Networking Site'$coefficients[['hispanic1']])-1)`
more likely to ask their connections on a social networking site about the 
Covid-19 vaccinations than non-Hispanics ( _p_ < 0.01).

Searching for information using an online search engine such as Google or Bing 
is theorically the search vehicle with the least amount of social or cognitive 
friction. Our descriptive statistics showed that not everyone used this vehicle,
with only 41% of our sample (See Table \@ref(tab:table-1-desc)) having sought
information about the Covid-19 vaccines using online search engines. My model
reveals interesting patterns to help explain the variation. As with previous 
models, receiving information of any kind increase the odds of online search, 
though to varying magnitudes (Doctor:
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_dr1']]), 2)` odds; 
Person:
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_pers1']]), 2)` odds; 
News or Television: 
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_tv1']]), 2)` odds; 
Social Networking Site:
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_onlnet1']]), 2)` odds; 
Online Forum: 
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_onlgroup1']]), 2)` odds).
I also find that older respondents were more likely to search online, with each 
addition year aged to yield
`r label_percent()(exp(part2_models$'Online Search'$coefficients[['age']])-1)`
higher odds of searching online ( _p_ < 0.001). The most interesting variations
come from what decreases the odds of online search. First, those with an
Associate's degree or higher were 
`r label_percent()(1-exp(part2_models$'Online Search'$coefficients[['college1']]))` 
less likely to search online for questions about Covid-19 than those with less
educational attainment ( _p_ < 0.001). In addition, I find that Hispanic-Latinos 
had `r label_percent()(1-exp(part2_models$'Online Search'$coefficients[['hispanic1']]))` 
lower odds of searching online than non-Hispanics ( _p_ < 0.05).

## Vaccination Views

```{r table-model-3}
#| results = "asis"

mod3 <- glm(formula = vacc_view_pos ~ receive_i_fr_dr +   
           receive_i_fr_pers +      
           receive_i_fr_tv +    
           receive_i_fr_onlnet +    
           receive_i_fr_onlgroup + 
           int_sought_fr_pers +  
           int_sought_fr_dr +    
           int_sought_fr_onlgroup +
           int_sought_fr_onlnet +
           int_sought_fr_onlsearch +
           age + 
           gender + 
           race_white +   
           race_black +    
           race_nat +      
           race_asian + hispanic +     
           college,  
         data = rain, family = binomial(link="logit"))

nullmod3 <- glm(formula = vacc_view_pos ~ 1,
            data = rain, family = binomial(link="logit"))
psuedor2_mod3 <- as.numeric(1-(logLik(mod3)/logLik(nullmod3)))

rows <- tribble(~term, ~model1,
                "$R_{McFadden}^2$", psuedor2_mod3)
attr(rows, 'position') <- c(19, 4)

modelsummary(mod3,
            gof_omit = 'AIC', statistic = NULL,
            estimate = "{estimate}{stars} ({std.error})",
            title = 'Does receiving or searching for information predict vaccination status?',
            notes = list('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Exponentiated Coefficients'),        
            coef_map = regression_labels,
            add_rows = rows,
            exponentiate = TRUE,
            output = "kableExtra") %>%
  kable_styling(latex_options = c("hold_position", "striped"))
```
Finally, in Table \@ref(tab:table-model-3), I investigate whether there is a
relationship between information receiving and information seeking behaviors and
the choice to get vaccinated against Covid-19 through a logistic regression
model predicting Vaccination Outlook. I find that receiving information from a
doctor or other medical profession is associated with 
`r round(exp(mod3$coefficients[['receive_i_fr_dr1']]), 2)` odds of receiving 
a Covid-19 vaccinated ( _p_ < 0.001) and that seeking information from a doctor
also increases the odds of vaccination by 
`r round(exp(mod3$coefficients[['int_sought_fr_dr1']]), 2)` ( _p_ < 0.001).
I find that my respondents who received information on a social networking site 
like Facebook or Twitter about the vaccines were 
`r label_percent()(1-round(exp(mod3$coefficients[['receive_i_fr_onlnet1']]), 2))` 
**less likely** to get vaccinated than those who received no information through
that vehicle ( _p_ < 0.001). However, I did find that those who reached out to
family or friends with their concerns were 
`r label_percent()(exp(mod3$coefficients[['int_sought_fr_pers1']])-1)` more likely
to become vaccinated ( _p_ < 0.05). Furthermore, my respondents who held an 
Associate's degree or above had `r round(exp(mod3$coefficients[['college1']]), 2)`
higher odds of becoming immunized ( _p_ < 0.001). And finally, I found further
ethno-racial differences in my sample. I found that Hispanic-Latinos had 
`r round(exp(mod3$coefficients[['hispanic1']]), 2)` higher odds of vaccination
than non-Hispanics ( _p_ < 0.05) and that my Asian sample had 
`r round(exp(mod3$coefficients[['race_asian1']]), 2)` higher odds as well over
non-Asians ( _p_ = 0.09). See Figure \@ref(fig:plot-model-3) for a visual 
representation of how these different coefficients are related to vaccination 
outlook. 
<!-- TODO: watch out about saying vaccinated, this variable is about opinion -->

```{r plot-model-3}
#| fig.cap='Plot of Coefficients, Model 3'

plot_summs(mod3,
           scale = TRUE,
           coefs =  c(
             'Received Info, Doctor' = 'receive_i_fr_dr',
             'Received Info, Person' = 'receive_i_fr_pers',
             'Received Info, News' = 'receive_i_fr_tv',
             'Received Info, Social Networking Site' = 'receive_i_fr_onlnet',
             'Received Info, Online Forum' = 'receive_i_fr_onlgroup',
             'Sought Info' = 'int_sought',
             'Sought Info, Doctor' = 'int_sought_fr_dr',
             'Sought Info, Person' = 'int_sought_fr_pers',
             'Sought Info, Social Networking Site' = 'int_sought_fr_onlnet',
             'Sought Info, Online Forum' = 'int_sought_fr_onlgroup',
             'Sought Info, Online Search' = 'int_sought_fr_onlsearch',
             'Age' = 'age',
             'Gender, Male' = 'gender=Male',
             'Gender, Other' = 'gender=Other',
             "Associate's Deg or above" = 'college',
             'White' = 'race_white',
             'Black' = 'race_black',
             'Native American' = 'race_nat',
             'Asian' = 'race_asian',
             'Hispanic or Latino/x' = 'hispanic'),
           robust = TRUE) +
  labs(title = 'Does receiving or searching for information predict vaccination sentiment?', 
       caption = 'Exponentiated Coefficients: Scaled and Robust') + 
  ylab("") +
  theme_diss()+
  scale_color_manual(values = my_palette[3])
```

# Discussion

summary of purpose

for results - Table 1
this never before seen pattern in sociology may be a result of response bias, (most likely to select more information) but it also could indicate a sheltering effect ...

rather than on “the receiver as an active information seeker and processor."
[@johnsonComprehensiveModelCancerRelated1993].

Est velit egestas dui id ornare arcu odio ut. Tempus iaculis urna id volutpat
lacus laoreet non curabitur gravida. Integer malesuada nunc vel risus commodo.
Ipsum nunc aliquet bibendum enim facilisis gravida. Ac turpis egestas maecenas
pharetra convallis posuere morbi leo. Id diam vel quam elementum pulvinar etiam.
Neque aliquam vestibulum morbi blandit cursus risus at ultrices mi. Sem
fringilla ut morbi tincidunt augue interdum velit euismod. Ullamcorper eget
nulla facilisi etiam dignissim diam quis enim lobortis. Turpis egestas pretium
aenean pharetra magna ac placerat vestibulum lectus. Elit ut aliquam purus sit
amet luctus venenatis. Vitae tempus quam pellentesque nec nam. In est ante in
nibh. Nisl tincidunt eget nullam non nisi est sit. Cras adipiscing enim eu
turpis egestas. Velit dignissim sodales ut eu sem integer vitae justo eget. Arcu
non sodales neque sodales ut etiam sit amet. Turpis in eu mi bibendum neque
egestas congue quisque egestas.

At varius vel pharetra vel turpis nunc eget lorem. In nulla posuere sollicitudin
aliquam. Dui faucibus in ornare quam. Vivamus arcu felis bibendum ut tristique.
Ornare aenean euismod elementum nisi. Eget aliquet nibh praesent tristique.
Rhoncus dolor purus non enim praesent elementum. Id faucibus nisl tincidunt eget
nullam non nisi est. Eget arcu dictum varius duis at consectetur lorem donec
massa. At tempor commodo ullamcorper a lacus vestibulum sed arcu.

Lectus mauris ultrices eros in. Diam ut venenatis tellus in metus vulputate eu
scelerisque. Magna ac placerat vestibulum lectus mauris. Est velit egestas dui
id ornare arcu odio ut. Sagittis aliquam malesuada bibendum arcu vitae
elementum. Nisi lacus sed viverra tellus in hac habitasse. Nunc eget lorem dolor
sed. Condimentum id venenatis a condimentum vitae. Ipsum faucibus vitae aliquet
nec ullamcorper sit amet risus nullam. Tincidunt eget nullam non nisi est.
Ornare quam viverra orci sagittis eu volutpat. Risus quis varius quam quisque.
Lobortis scelerisque fermentum dui faucibus. A diam sollicitudin tempor id eu
nisl nunc mi ipsum.

\newpage
# References {-}

<div id="refs"></div>

\newpage
# Appendix 1: Survey Questions

-------------------------------------------------

*Q181*: Now, we're going to transition a bit to ask you about the information you received during the Covid-19 pandemic.

-------------------------------------------------

*Q170*: In the past 12 months, without searching for it, did you receive information about the Covid-19 vaccine from … (check all sources you received information from)

*	your doctor or other health professional?   (1) 
*	a person like friend, neighbor, or family member that you know ?  (9) 
*	From television news channel or the newspaper?  (10) 
*	From an online discussion group, listserve, or other online forum including a Facebook group or subreddit?  (11) 
*	From a social networking site such as Facebook timeline, Twitter status update, or linkedin?  (12) 
*	Other  (13) _______

-------------------------------------------------

*Q171*: And how about you yourself intentionally looking for information about a Covid-19 vaccine? Such information can include things such as advice, clarification, facts, and experiences.

*	Yes  (4) 
*	No  (5)

-------------------------------------------------

*Q172*: How did you look for information about the Covid-19 vaccine?

*	Asking a person like friend, neighbor, or family member that I know  (1) 
*	Asking my doctor or another health professional  (4) 
*	Posted queries in an online discussion group, listserve, or other online forum like a Facebook Group or Subreddit  (5) 
*	Posted queries on a social networking site such as Facebook timeline, Twitter status update, or Linkedin  (6) 
*	Searched for my question using an online search engine such as Google or Bing  (7) 
*	Other  (8) ___________

-------------------------------------------------

*Q173*: What sort of information did you search for? 
 Separate different topics with a comma or semi-colon.

-------------------------------------------------

*Q174*: How useful was the information you found?

*	Extremely useful  (22) 
*	Very useful  (23) 
*	Moderately useful  (24) 
*	Slightly useful  (25) 
*	Not at all useful  (26)

-------------------------------------------------

*Q175*: Did the information you learned affect your decision to get vaccinated against Covid-19?

*	Yes  (39) 
*	No  (40)

-------------------------------------------------

*Q179*: Did you receive a Covid-19 vaccine?

*	Yes  (9) 
*	No  (10) 
*	I'm unsure or would not like to respond  (11)

-------------------------------------------------

*Q180*: Do you plan to receive a vaccine for the prevention of the Covid-19 virus?

*	Definitely not  (9) 
*	Probably not  (10) 
*	Might or might not  (11) 
*	Probably yes  (12) 
*	Definitely yes  (13) 
*	I would not like to respond  (14)

-------------------------------------------------

*gender*: What is your gender?

*	Male  (1) 
*	Female  (2) 
*	Other  (3) ________________
*	Prefer not to say  (4)

-------------------------------------------------

*hispanic*: Are you Hispanic, Latino/a/x, or Latin American Origin?

*	Yes  (1) 
*	No  (2)

-------------------------------------------------

*race*: What is your race? If you are "mixed race", select all that apply.

*	White  (1) 
*	Black or African American  (2) 
*	American Indian or Alaskan Native  (3) 
*	Asian (please specify):  (4) _______
*	Other (please specify):  (5) _______
*	Prefer not to say  (6)

-------------------------------------------------

*educ*: What is the highest level of education that you have completed?

*	Less than high school  (1) 
*	High school graduate  (2) 
*	Some college  (3) 
*	Associate's or Technical degree  (4) 
*	Bachelor's degree  (5) 
*	Graduate or professional degree  (6)

-------------------------------------------------

*income*: We would be interested to know roughly what your total household income
before taxes is. We mean income from all sources, including welfare, stock
dividends, other household members' income, etc. In 2020, in which bracket did your total family income fall?

*	Under $1,000  (1) 
*	\$1,000 to \$9,999   (2) 
*	\$10,000 to \$19,999  (3) 
*	\$20,000 to \$29,999  (4) 
*	\$30,000 to \$39,999  (5) 
*	\$40,000 to \$49,999  (6) 
*	\$50,000 to \$59,999  (7) 
*	\$60,000 to \$74,999  (8) 
*	\$75,000 to \$89,999   (9) 
*	\$90,000 to \$109,999  (10) 
*	\$110,000 to \$129,999  (11) 
*	\$130,000 to \$149,999  (12) 
*	\$150,000 or over  (13) 
*	Don't know  (14)

-------------------------------------------------

*age*: What is your age? (in years)