### 2022-01-24 paper2.Rmd
adding to front half

---
title: "Article 2:	Information-Seeking Behavior Decision-Making between Online Search and Interpersonal Networks: The Case of Covid-19 Vaccinations "
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output:
  bookdown::pdf_document2: 
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "My Library.bib"
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      cache = TRUE,
                      fig.width = 7,
                      fig.asp = 0.8,
                      out.width = "80%",
                      fig.align="center"
                      # dev.args = list(png = list(type = "cairo"))
                      )

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales)

rain <- read_rds(here('data', 'RAiN_vacc_clean.rds')) %>% 
  mutate(across(where(is.logical), as.factor),
         date = lubridate::as_date(EndDate))

my_palette = c('#56b4e9','#999999','#E69F00','#009E73','#CC79A7','#0072B2','#D55E00','#E69F00')
my_palette  <- c(MetBrewer::met.brewer(name = 'Cross', type = 'discrete'), '#555F61')

## Automatically use showtext to render text
font_add_google("Lora", "lora")
showtext_auto()

theme_diss <- function(base_size = 14) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      # Figure assembly
      plot.title = element_text(family="lora",
                                size = rel(1), 
                                margin = margin(0,0,5,0), 
                                hjust = 0),
      plot.title.position = "plot",
      plot.subtitle = element_text(family="lora",
                                   size = rel(0.85)),
      plot.caption = element_text(family="lora",
                                  size = rel(0.70),
                                  hjust = 1),
      # Graphical Zones
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      # Axes
      axis.title = element_text(family="lora", 
                                size = rel(0.85)),
      axis.text = element_text(family="lora", 
                               size = rel(0.70)),
      axis.line = element_line(color = "black", 
                               arrow = arrow(length = unit(0.3, "lines"), 
                                             type = "closed")),
      # Legend
      legend.title = element_text(family="lora",
                                  size = rel(0.85)),
      legend.text = element_text(family="lora", 
                                 size = rel(0.70)),
      legend.key = element_rect(fill = "transparent", 
                                colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", 
                                       colour = NA),
      # Facetting 
      strip.background = element_rect(fill = "#17252D",
                                      color = "#17252D"),
      strip.text = element_text(family="lora",
                                size = rel(0.85), 
                                color = "white", 
                                margin = margin(5,0,5,0))
    )
}

theme_set(theme_diss())

conf_int_prop <- function(var){
  p <- mean(!!var, na.rm = T)
  q <- 1 - p
  n <- nrow(rain)
  min <- p - 1.96*sqrt((p*q)/n)
  max <- p + 1.96*sqrt((p*q)/n) 
  paste0(round(min, 2), ", ", round(max, 2))
}
print_perc <- function(var){
  label_percent()(mean(var, na.rm = T))
}

```

# Abstract 
How do individuals perform information search, and what factors help them decide
which information search vehicle to utilize, such as interpersonal networks or
online tools, I investigate the above question using individual survey data of
934 Adults in the United States and find that X and Y. Furthermore, this paper
demonstrates that searching for a question online using a search engine or
social network was more common than traditional means of information gathering,
such as network activation (asking someone in their social network for
informational support). Using X and Y theory I predict that that few people will
activate their social network for informational support in my chosen case of
Covid-19 vaccinations because of the social cost of doing so but fail to find
evidence for this thesis. I also investigate which information channels are
associated with vaccinations against the virus, and find that social netwokring
sites are associated with lower rates. I also find important ethnocracial
differences in information search vehicles utilization and vaccination
reception.

<!-- TODO this needs a lot of work but its a start -->
Previous research has greatly failed to distinguish between the activation of
information seeking behaviors online and offline. In this paper, I aim to
uncover just how different various methods of information search are and begin a
line of research inquiry that investigates the factors associated with each vehicle. 
Finally, I also aimed to uncover how the different fountains of information and
information search vehicles hold real world consequences through their associations
with Covid-19 vaccination rates and intentions.

# Intro

Information is all around us and permeates every facet of human society, from
the hordes of data gathered during the internet age or passing gossip between
friends. Human communication and human society are based on the circulation of
information and knowledge. Information can be shared by others through
information campaigns and consumed by an individual, called information push, or
intentionally sought out by individuals, called information pull 
[@cybenkoFoundationsInformationPush1999].

Information is especially important given the prominent trends of
misinformation, disinformation, and mistrust in traditional institutions
[@starbird19; @kata10]. While no belief in false information can be thought of as
harmless [@douglas21] some discredited theories such as 'vaccines cause autism'
have major ramifications for both the individual and the social world that
surrounds them. For example, a new found public resistance to vaccination
against Measles by so-called 'anti-vaxxers' caused a 17% increase in rates
worldwide in 2019. This increase killed  142,000 people, most of whom were
children under the age of five [@givetash19]. However, even in the ever-evolving
research agenda of misinformation, much of the research focuses on spaces where
individual first come into contact with conspiracy theories or other information
rather than what individuals do with the information they obtain
[@johnsonOnlineCompetitionPro2020; @broniatowski_etal20]. 
<!-- cite something else here -->

The lack of research in this area makes it quite unclear how individuals choose
to search for information through the multimedia environment of the 21st
century. Even outside of the misinformation literature, much of the research
conducted in social networks and communication focus on senders, influencers,
and persuasion strategies [@mertonManifestLatentFunctions1968;
@katzPersonalInfluencePart1955; @lazarsfeldPeopleChoice1944 ], rather than on
“the receiver as an active information seeker and processor."   [@johnsonComprehensiveModelCancerRelated1993].

<!-- TODO cleanup -->
For this paper, I will use the case of Covid-19 vaccinations to explore the
variation among information search vehicles. T

Research Questions

* How do computer-mediated or interpersonal network information-seeking strategies vary across populations? 
* What factors lead people to perform health information search online versus among social network ties? 
* How does the utilization of social networks, search engines, and social media as vary in the case of searching for information on the Covid-19 vaccine?

<!-- TODO  cleanup -->

## subheading - how people search

Active directed searching by individuals to obtain information is sparsely
discussed in the sociological literature. Pejtersen
[-@pejtersenDesignComputeraidedUsersystem1984], a scholar of library and
information science, theorized that there are 5 strategies for searching for
information. The most common strategy is browsing, where people follow leads
based on associations without planning ahead. Another strategy is analytical,
which includes an explicit consideration of all facets of the question to guide
a search. The empirical method guides the search based on tactics that were
successful in past research. The known site strategy is to go to the direct
source of the information if known. And finally, the similarity method is to
find information based on another similar question that already has an answer.
These 5 strategies vary in their demands for prior knowledge, cognitive
processing, memory, and time spent. While this theory is aimed at finding
information in a library setting, scholars have extended the theory to other
fields and validated the framework [@fidelHumanInformationInteraction2012]; the
frame is a useful beginning point for my theories of information search through
network activation or through computer-mediated communication.

Some communication theorists ask why an answer to a question is sought in the
first place. The Theory of Uncertainty Management
[@brashersCommunicationUncertaintyManagement2001] professes that people search
for information when their uncertainty around the subject leads to anxiety or
other cognitive harms. The Theory of Motivated Information Management
[@afifiTheoryMotivatedInformation2004; @afifiSeekingInformationSexual2006]
extends the prior by adding that uncertainty itself is not the catalyst for
information-search; rather, it is driven by a discrepancy between the current
level of uncertainty on a subject and desired level of uncertainty.

One way to find information is to activate network ties to find out information
through a form of social support. Social support, while previously used
interchangeably with the term social networks and social integration
[@houseStructuresProcessesSocial1988], are the emotional, informational, and
instrumental assistance functions performed between social ties and have strong
and measurable association with health outcomes
[@houseMeasuresConceptsSocial1985; @thoitsMechanismsLinkingSocial2011].
Informational support is the process of seeking “help in defining,
understanding, and coping with problematic events and include education, advice,
or referral to another source of support” [@winemiller_etal93, p. 640].
Brashers, a health communications researcher, defines informational support
slightly differently, focusing on the exchange of information that “facilitates
coping with life stresses… that may be exchanged among members of a support
network” [-@brashersInformationSeekingAvoiding2002, p. 260] <!--  Both of these
definitions provide important lenses for my question of how computer-mediated or
interpersonal information-seeking strategies vary.--> If information is needed
but there is no search for coping or deeper understanding of important matters,
people are likely to choose to search online.

Core discussion networks are the default lens to investigate network resources
and social support in surveys. The underlying assumption is that individuals
reach out to a handful of strong ties when in need of support, which can be
elicited in surveys using name generators [@marsdenCoreDiscussionNetworks1987].
This approach has yielded important insights, but largely overlooks crucial
processes of resource activation [@hurlbertCoreNetworksTie2000;
@perrySocialNetworkActivation2015; @smithDonPutMy2005]. Small
[-@smallSomeoneTalk2017] shows that the core discussion network does not capture
how people activate social support in practice and indicates that people draw on
much broader social connections for support, also calling back to the weak ties
research by Granovetter [-@granovetterStrengthWeakTies1973]. There is ongoing
research investigating how weak-tie support holds for different support types
based on the architecture put forward by House, Umberson, and Landis
[-@houseStructuresProcessesSocial1988] of instrumental, emotional, and
informational support. I follow the definition of Cohen and Wills
[-@cohenStressSocialSupport1985] and define informational support as “help in
defining, understanding,  and coping with problematic events and include
education, advice, or referral to another source of  support”
[@winemiller_etal93, p. 640].

However, informational support can also be sought outside of the social network
context, namely via computer-mediated information search tools such as the
process of “Googling”. As the online environment began penetrating all facets of
modern human life, it makes sense that performing online search has become one
of the most convenient vehicles for information search. While Small
[-@smallSomeoneTalk2017] focuses on how support can depend on a network tie
simply "happened to be there", online search is theoretically the most
frictionless and costless mode of support. There are three computer-mediated
vehicles I consider for this analysis: search engines like Google, posting
questions on online forums like subreddits or Facebook groups, and posting a
status update online via Facebook or Twitter. The first two comprise what I call
“online search”, while the third is conceptually distinct and important to
include in the survey; yet, it is a hybrid of social network activation and
online search.

From studies of mass communication, uses and gratifications theory (UGT)
[@blumlerUsesMassCommunications1974; @tanMassCommunicationTheories1985] may also
lend itself to theorizing why computer-mediated or interpersonal
information-seeking strategies vary. UGT posits that users are not passive
consumers of media and that people have an active role in choosing different
sources of media based on their satisfaction of specific needs on an individual
basis. UGT is based on Maslow’s [-@maslowTheoryHumanMotivation1943] hierarchy of
needs and is compatible with Lazarsfeld and Katz’s theory of two-step flow
[-@katzPersonalInfluencePart1955] because people can choose their media and the
opinion leaders they follow. Modern-day theorists have extended UGT theory and
classified the uses and gratifications of the internet and of social media.
Stafford, Stafford, and Schkade [-@staffordDeterminingUsesGratifications2004]
theorize that the internet provides gratification through useful content that
meets expectations, gratification from purposeful navigating or random browsing
as a process, and social gratification from forming and deepening social ties.
Leung [-@leungGenerationalDifferencesContent2013] theorizes that social media is
gratifying for users because it allows for venting of negative feelings,
provides recognition, provides entertainment, promotes social affection, and
fulfills cognitive needs. Adapting UGT to my own purposes, I theorize that
informational support will be activated from network ties when the informational
need is related to forming and deepening social ties but search will be
conducted online when information is needed but an individual does not have an
additional cognitive need to fill through social interaction. Alternative goals
such as identity management or relational maintenance
[@brashersInformationSeekingAvoiding2002] need to be balanced and will determine
where search is conducted.

<!-- connect above and below paragraph to flow together -->
<!-- TODO  cleanup -->

As online search or interaction also allows increased anonymity, no or lowered
social cost, and the potential avoidance of embarrassment and other negative
social interactions, I theorize that a majority of people who searched for
information will have done so online.

there are theoretical reasons why individuals would choose to search for
information among their network ties instead of online search. For instance,
Rains [-@rainsCopingIllnessDigitally2018] finds that patients tend to search for
technical information about an illness online but turn to their social network
for experiential information from others facing similar circumstances in the
case of cancer. Moreover, if an individual distrusts the medical establishment,
or popular media, they may be more likely to activate informational support
among their social network or turn to online groups that validate their
worldview [@bogersHowSocialAre2014].

<!-- TODO add a hypothesis here -->

# Research Methodology
The data used for this research project are based on original survey data
sampled between `r format( min(rain$date), '%B %d, %Y')`  through 
`r format(max(rain$date), '%B %d, %Y')`. This survey was hosted on Qualtrics and
participants were paid and recruited using Amazon Turk (MTurk). The total valid
survey responses (_n_ = `r nrow(rain)`) were selected from a total of 1,066
respondents; some responses were disqualified due to a few factors such as
detected usage of a VPN, random answer clicking (determined by illogical
responses), poor quality typed responses (e.g. social network alters
consistently named random nouns), or taking the survey more than once. The same
is slighted gender unbalanced, slightly skewing towards a male sample. For the
respondents who provided their zipcode (p = 
`r round(nrow(rain[!is.na(rain$region),]) / nrow(rain), 2)`), the 
sample is quite balanced with a slight skew towards MTurkers in the South (p = 
`r round(nrow(rain[rain$region == 'South',]) / nrow(rain), 2)`). 
The survey took an average of `r round(mean(rain$duration), 2)` minutes with a 
standard deviation of `r round(sd(rain$duration), 2)` minutes. The shortest 
valid survey took `r round(min(rain$duration), 2)` minutes and the longest took 
`r round(max(rain$duration), 2)` minutes. Participants were paid \$6.00 for their 
time after Amazon administrative fees, funded by a Grant awarded to Kelsey E.
Gonzalez and Nicolas Legewie by the Summer Institute in Computational Social
Science and the Russell Sage Foundation. The main portion of this survey
replicates the book, Someone to Talk To [@smallSomeoneTalk2017] by expanding on
Small’s finding that people draw on much broader sources than “important people”
for support by segregating weak-tie social support into instrumental, emotional,
and informational support (see House et al. 1988)). The latter part of the
survey focuses on resource activation for the specific case-study of information
during the Covid-19 pandemic and is analyzed in this paper.

```{r table-1-desc}
#| results = "asis"

desc_table <- rain %>% 
  select(receive_i_fr_dr,
  receive_i_fr_pers,
  receive_i_fr_tv,
  receive_i_fr_onlnet,
  receive_i_fr_onlgroup,
  int_sought,
  int_sought_fr_pers,
  int_sought_fr_dr,
  int_sought_fr_onlgroup,
  int_sought_fr_onlnet,
  int_sought_fr_onlsearch,
  int_sought_affect,
  vacc_status,
  vacc_view_pos,
  race_white,
  race_black,
  race_nat,
  race_asian, 
  hispanic,
  college,
  age, receive_i) %>% 
  as.data.frame()

datasummary(formula = 
  (`Received info` = receive_i) +
    (`Received info, Doctor` = receive_i_fr_dr) +
    (`Received info, Person` = receive_i_fr_pers) +
    (`Received info, News` = receive_i_fr_tv) +
    (`Received info, Social Networking Site` = receive_i_fr_onlnet) +
    (`Received info, Online Forum` = receive_i_fr_onlgroup) +
    (`Sought Info` = int_sought) +
    (`Sought Info, Doctor` = int_sought_fr_dr) +
    (`Sought Info, Person` = int_sought_fr_pers) +
    (`Sought Info, Social Networking Site` = int_sought_fr_onlnet) +
    (`Sought Info, Online Forum` = int_sought_fr_onlgroup) +
    (`Sought Info, Online Search` = int_sought_fr_onlsearch) +
    (`Vaccination Status` = vacc_status) +
    (`Vaccination Favorability` = vacc_view_pos) +
    (`Age` = age) +
    (`Hispanic or Latino/x` = hispanic) +
    (`Race, Black` = race_white) +
    (`Race, Black` = race_black) +
    (`Race, Native American` = race_nat) +
    (`Race, Asian or Pacific Islander` = race_asian) +
    (`Associate's Deg or above` = college)  ~ 
    (Mean + SD) ,
  data = desc_table,
  fmt = '%.2f',
  title = 'Descriptive Statistics for Dichotomous and Numeric Variables',
  notes =c(glue("Notes: {nrow(rain)} Surveyed, Conducted {format( min(rain$date), '%B %d')} through {format(max(rain$date), '%B %d')}, 2021.")),
  output = "kableExtra") %>% 
  kable_styling(latex_options = c("hold_position", "striped"))

```

## Measures

###  Sought Information
There are multiple dependent variables of interest for this paper. The first
variable is a dichotomous indicator of whether someone intentionally sought out
information about Covid-19 vaccinations. The survey asked, "And how about you 
yourself intentionally looking for information about a Covid-19 vaccine? Such 
information can include things such as advice, clarification, facts, and
experiences". About `r print_perc(rain$int_sought)` ($\widehat{p}$)
of respondents had intentionally sought out information about Covid-19 
vaccinations (95% Confidence Interval = \{`r conf_int_prop(rain$int_sought)`\}).

Further dependent variables segment the above question into multiple options;
"How did you look for information about the Covid-19 vaccine?". Respondents were
posed with 5 responses and 'other', of which they could select multiple. The
largest proportion, `r print_perc(rain$int_sought_fr_dr)`, asked their 
'doctor or another health professional' (95% Confidence Interval = 
\{`r conf_int_prop(rain$int_sought_fr_dr)`\}).
About `r print_perc(rain$int_sought_fr_pers)`  said they asked 'a person 
like friend, neighbor, or family member that [they] know' (95% Confidence 
Interval = \{`r conf_int_prop(rain$int_sought_fr_pers)`\}). An additional
`r print_perc(rain$int_sought_fr_onlsearch)` searched 'for [their] question 
using an online search engine such as Google or Bing' (95% Confidence Interval 
= \{`r conf_int_prop(rain$int_sought_fr_onlsearch)`\}). About a quarter of the 
sample, `r print_perc(rain$int_sought_fr_onlgroup)`, 'posted queries in an 
online discussion group, listserve, or other online forum like a Facebook Group 
or Subreddit' (95% Confidence Interval = \{`r conf_int_prop(rain$int_sought_fr_onlgroup)`\}).
And finally, `r print_perc(rain$int_sought_fr_onlnet)` of the sample 'posted queries 
on a social networking site such as Facebook timeline, Twitter status update, or 
Linkedin' (95% Confidence Interval = \{`r conf_int_prop(rain$int_sought_fr_onlnet)`\}).
Much of the question wording and the specific categories of search were inspired
by ICPSR project 37220 [@scanlon19].

```{r table-2-desc}
#| results = "asis"

datasummary(formula = (
  # (`Income` = income)+  
                         (`Gender` = gender)+  
                         (`Highest Education Level` = educ) + 
                         (`Plan to be Vaccinated if not` = vacc_status_future)  +
                         (`Found Info Sought Useful` = int_sought_useful) +
                         (`Census Region` = region)) ~ N + Percent(), 
 data = rain,
   fmt = '%.2f',
  title = 'Descriptive Statistics for Categorical Variables',
 notes =c(glue("Notes: {nrow(rain)} Surveyed, Conducted {format( min(rain$date), '%B %d')} through {format(max(rain$date), '%B %d')}, 2021.")),
 output = "kableExtra") %>%
  kable_styling(latex_options = c("hold_position", "striped"))

# TODO reorder factors in categorical table

```

### Vaccination Outlook

```{r}
vacc_status_freq <- rain[!is.na(rain$vacc_status_future),] %>% 
  group_by(col = vacc_status_future) %>% 
  summarise(n = n()) %>%
  mutate(n = n / sum(n))
```
Another Dependent variable for this analysis is a dichotomous variable
indicating whether the respondent has received a vaccination against Covid-19 or
plans to. For this survey, the number of doses were unimportant and actually
having already received the dose were irrelevant to the research question. This
variable instead focuses on intent or opinion about the vaccine. About 
`r print_perc(rain$vacc_view_pos)` have a positive view about the vaccine. This
construct is made through the combination of  two survey items. The first item
is actual vaccination status, obtained through the question 'Did you receive a
Covid-19 vaccine?'. `r print_perc(rain$vacc_status)` of the sample said that
they had indeed received a vaccination (95% Confidence Interval = \{
`r conf_int_prop(rain$vacc_status)`\}). As the national vaccination rate is
closer to 55% (CITE), this indicates some bias in our sample that must be
acknowledged: either MTurkers are more likely to be vaccinated that the normal
population (sampling bias) or there is some conformity bias in the responses
causing survey takers to provide false answers. The second survey item that
contributes to 'Vaccination Outlook' was only given to those who had not
responded 'yes' to the to their vaccination status. These respondents were asked 'Do you
plan to receive a vaccine for the prevention of the Covid-19 virus?'. Of those
who were asked this question (n = `r nrow(rain[!is.na(rain$vacc_status_future),])`),
`r label_percent()(filter(vacc_status_freq, col == 'Definitely yes') %>% pull(n))`
said they were 'definitely yes' getting the vaccine; 
`r label_percent()(filter(vacc_status_freq, col == 'Probably yes') %>% pull(n))` 
said they were 'probably yes' going to receive vaccine. The rest of the sample was more unsure: 
`r label_percent()(filter(vacc_status_freq, col == 'Might or might not') %>% pull(n))`
said they 'Might or might not', 
`r label_percent()(filter(vacc_status_freq, col == 'Probably not') %>% pull(n))`
said 'Probably not', and finally
`r label_percent()(filter(vacc_status_freq, col == 'Definitely not') %>% pull(n))`
said they would 'definitely not' receive the vaccine. The variable 'Vaccination
Outlook' indicates their either the respondent received a Covid-19 vaccine or
either probably or definitely will in the future.

### Independent Variables
The first independent variables used in this analysis are based on the question, 
'In the past 12 months, without searching for it, did you receive information
about the Covid-19 vaccine from … (check all sources you received information
from)'. This differs from the 'How did you look for information' question above
because it is based on passive reception of information. The largest proportion
of respondents had received information from a person like friend, neighbor, or
family member that [they] know ($\widehat{p}$ = `r print_perc(rain$receive_i_fr_pers)`
). Respondents also commonly received information from a television news channel
or a newspaper, about `r print_perc(rain$receive_i_fr_tv)` of the sample.
`r print_perc(rain$receive_i_fr_dr)` of the sample had received information from
their doctor or other health professional while `r print_perc(rain$receive_i_fr_onlnet)` 
received information from a social networking site such as Facebook timeline,
Twitter status update, or linkedin. Finally, the lowest proportion of the sample
had received information from an online discussion group, listserve, or other
online forum like a Facebook group or subreddit 
($\widehat{p}$ = `r print_perc(rain$receive_i_fr_onlgroup)`).

We also asked various demographic questions to understand our sample. The 
average age of our sample was `r round(mean(rain$age, na.rm = T), 2)`. 
Our sample was `r label_percent()(nrow(filter(rain, gender == "Female")) / nrow(rain))`
female, and `r label_percent()(nrow(filter(rain, gender == "Male")) / nrow(rain))`
male. The sample is diverse racially though some ethnic groups are under
represented compared to national averages. Race was evaluated using the
question, 'What is your race? If you are "mixed race", select all that apply.'
Because of the multi-selection question, each variable is dichotomous and
proportions are shown. `r print_perc(rain$race_white)` of the sample claimed
they were 'White', `r print_perc(rain$race_black)` claimed to be 'Black or
African American', `r print_perc(rain$race_nat)` chose 'American Indian or
Alaskan Native', `r print_perc(rain$race_asian)` selected 'Asian', and finally 
when asked ' Are you Hispanic, Latino/a/x, or Latin American Origin?', 
`r print_perc(rain$hispanic)` selected 'Yes'. Respondents were also asked to 
select the  highest level of education that you have completed. Based on the 
breakdown between 'Less than high school', 'High school graduate', 'Some college',
'Associate's or Technical degree', 'Bachelor's degree', and 'Graduate or 
professional degree',  `r print_perc(rain$college)` of the sample were classified
to have a College-level education (Associate's degree or above).

# Analysis

My analytic strategy proceeds in four steps. In Tables \@ref(tab:table-1-desc) 
& \@ref(tab:table-2-desc), I present descriptive statistics for all study 
variables, including variable ranges, means, and standard deviations. For this 
paper, I rely heavily on logistic regression, modeled in `r`; specifically, I 
fit a series of Binomial Generalized Linear Models [@venables2002a]. In Table \@ref(tab:table-model-1)^[All tables in this paper were created with the modelsummary package in `r` [@modelsummary] ]
, I fit a logistic regression model predicting
whether someone sought information about Covid-19 using their sources of
information ('Received info') as well as various demographic variables such as
age, education level and race. Figure \@ref(fig:plot-model-1) then illustrates
the relationship between the coefficients and information seeking.^[Figures in 
this paper were all created using the `r` packages jtools [@jtools] and ggplot2 [@wickham_etal; @wickham11] ].
In Table \@ref(tab:model-2-phi) I disaggregate information seeking behaviors by
investigating the difference in proportions of each of the 5 different avenues
of seeking information about Covid-19: from another person, from a doctor, on an
online forum, on a social networking site, or using an online search tool. I
investigate these differences through a phi coefficient [@warrens08], otherwise
known as a mean square contingency coefficient which is used to investigate the
degree of association between two binary variables. ^[To calculate the Phi
Coefficient, I utilize the Psych `r` package [@psych] ] I then investiage the
associations between my predictors and the 5 information search vehicles by
modelling a series of logistic regressions in Table \@ref(tab:table-model-2).
Finally, in Table \@ref(tab:table-model-3), I investigate whether there is a
relationship between information receiving and information seeking behaviors and
the choice to get vaccinated against Covid-19 through a logistic regression
model predicting Vaccination Outlook.  Figure \@ref(fig:plot-model-3) then
illustrates the relationship between the coefficients and vaccination outlook.

# Results
```{r clean_before_reg}
rain <- rain %>% 
  mutate(across(c(receive_i,
                  receive_i_fr_dr,
                  receive_i_fr_pers,
                  receive_i_fr_tv,
                  receive_i_fr_onlnet,
                  receive_i_fr_onlgroup,
                  int_sought,
                  int_sought_fr_pers,
                  int_sought_fr_dr,
                  int_sought_fr_onlgroup,
                  int_sought_fr_onlnet,
                  int_sought_fr_onlsearch,
                  int_sought_affect,
                  vacc_status,
                  vacc_view_pos,
                  hispanic,
                  race_white,
                  race_black,
                  race_nat,
                  race_asian, 
                  college), ~ factor(.x, levels = c(0, 1))))

regression_labels = c(
  'receive_i_fr_dr1' = 'Received Info, Doctor',
  'receive_i_fr_pers1' = 'Received Info, Person',
  'receive_i_fr_tv1' = 'Received Info, News',
  'receive_i_fr_onlnet1' = 'Received Info, Social Networking Site',
  'receive_i_fr_onlgroup1' = 'Received Info, Online Forum',
  'int_sought1' = 'Sought Info',
  'int_sought_fr_dr1' = 'Sought Info, Doctor',
  'int_sought_fr_pers1' = 'Sought Info, Person',
  'int_sought_fr_onlnet1' = 'Sought Info, Social Networking Site',
  'int_sought_fr_onlgroup1' = 'Sought Info, Online Forum',
  'int_sought_fr_onlsearch1' = 'Sought Info, Online Search',
  'age' = 'Age',
  'gender=Male' = 'Gender, Male',
  'gender=Other' = 'Gender, Other',
  'college1' = "Associate's Deg or above",
  'race_white1' = 'White',
  'race_black1' = 'Black',
  'race_nat1' = 'Native American',
  'race_asian1' = 'Asian',
  'hispanic1' = 'Hispanic or Latino/x')
```
## Information Seeking
```{r table-model-1}
#| results = "asis"

mod1 <- glm(formula = int_sought ~ 
              receive_i_fr_dr +   
              receive_i_fr_pers +   
              receive_i_fr_tv +   
              receive_i_fr_onlnet +   
              receive_i_fr_onlgroup +   
              age + 
              gender + 
              race_white +   
              race_black +    
              race_nat +      
              race_asian +  
              hispanic  + 
              college,
            data = rain, family = binomial(link="logit"))
nullmod1 <- glm(formula = int_sought ~ 1,
            data = rain, family = binomial(link="logit"))
psuedor2 <- as.numeric(1-(logLik(mod1)/logLik(nullmod1)))

rows <- tribble(~term, ~model1,
                "$R_{McFadden}^2$", psuedor2)
attr(rows, 'position') <- c(14, 4)

modelsummary(mod1,
            gof_omit = 'AIC', statistic = NULL,
            estimate = "{estimate}{stars} ({std.error})",
            title = 'What predicts someone intentionally searched for information',
            add_rows = rows,
            coef_map = regression_labels,
            notes = list('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Raw, Unexponentiated Coefficients'),            
            # exponentiate = TRUE,
            output = "kableExtra") %>%
  kable_styling(latex_options = c("hold_position", "striped"))

```

```{r plot-model-1}
#| fig.cap='Plot of Coefficients, Model 1'

plot_summs(mod1, 
           scale = TRUE,
           coefs =  c(
             'Received Info, Doctor' = 'receive_i_fr_dr',
             'Received Info, Person' = 'receive_i_fr_pers',
             'Received Info, News' = 'receive_i_fr_tv',
             'Received Info, Social Networking Site' = 'receive_i_fr_onlnet',
             'Received Info, Online Forum' = 'receive_i_fr_onlgroup',
             'Sought Info' = 'int_sought',
             'Sought Info, Doctor' = 'int_sought_fr_dr',
             'Sought Info, Person' = 'int_sought_fr_pers',
             'Sought Info, Social Networking Site' = 'int_sought_fr_onlnet',
             'Sought Info, Online Forum' = 'int_sought_fr_onlgroup',
             'Sought Info, Online Search' = 'int_sought_fr_onlsearch',
             'Age' = 'age',
             'Gender, Male' = 'gender=Male',
             'Gender, Other' = 'gender=Other',
             "Associate's Deg or above" = 'college',
             'White' = 'race_white',
             'Black' = 'race_black',
             'Native American' = 'race_nat',
             'Asian' = 'race_asian',
             'Hispanic or Latino/x' = 'hispanic'),
           robust = TRUE) +
  labs(title = 'What predicts someone intentionally searched for information',
       caption = 'Exponentiated Coefficients: Scaled and Robust') +
  ylab("") +
  theme_diss() +
  scale_color_manual(values = my_palette[3])
```

Table \@ref(tab:table-model-1) provides the results of my logistic regression
analysis predicting whether someone sought information about Covid-19. These
initial results suggest similar results for each vehicle of information
delivery: each mode of receiving information whether from a doctor, a person,
the news, a social networking site, or an online forum, all increase the odds of
seeking out information. For example, respondents who had received information 
from a doctor or other medical professional were `r round(exp(mod1$coefficients[['receive_i_fr_dr1']]),2)`
(relative odds ratio = $exp(\beta k)$, _p_ < 0.001) times more likely to seek
out information over someone who hadn't. Similarly, respondents who had received 
information from 'a person  like friend, neighbor, or family member that [they] know'
were `r round(exp(mod1$coefficients[['receive_i_fr_pers1']]),2)`
( _p_ < 0.001) times more likely to seek out information over someone who hadn't
received information from a person like that. Interesting, those who had
received information from an online discussion group, listserve, or other online
forum including a Facebook group or subreddit were 
`r round(exp(mod1$coefficients[['receive_i_fr_onlgroup1']]), 2)` ( _p_ < 0.001)
times more likely to search out more information than someone who hadn't. While 
most demographic variables don't seem to have a strong effect on whether someone
sought out information about the Covid-19 vaccine, a few key variables had strong
associations. Age, for instance, is associated with seeking information; every 
additional year of age a person holds they are
`r label_percent()(exp(mod1$coefficients[['age']])-1)` 
more likely to search for information, possibly indicating a generational effect
of information seeking patterns ( _p_ < 0.01). Additionally, those who were more 
educated, that is those with an associate's degree or higher, were 
`r round(exp(mod1$coefficients[['college1']]), 2)` 
more likely to seek additional information than those without the degree 
( _p_ < 0.05). And finally, there is evidence to suggest that the Black or 
African American sample was more likely to seek out information, specifically: 
Blacks were `r round(exp(mod1$coefficients[['race_black1']]), 2)` times more likely 
to seek out information over non-Blacks ( _p_ < 0.05). Figure \@ref(fig:plot-model-1) illustrates
the relationship between the coefficients and information seeking.

## Information Search Vehicles

```{r model-2-phi}
phi <- rain %>% 
  select(pers = int_sought_fr_pers,
         dr = int_sought_fr_dr,
         forum = int_sought_fr_onlgroup,
         net = int_sought_fr_onlnet,
         search = int_sought_fr_onlsearch)

find_phi <- function(x, y){
  tab <- table(x, y)
  phi(tab, digits = 3)
}

phi_squared_tests <- 
  tribble(~col, ~Person, ~Dr, ~`Online Forum`, ~`Social Networking Site`, ~`Online Search`,
          'Person',   NA_complex_, find_phi(phi$pers, phi$dr), find_phi(phi$pers, phi$forum), find_phi(phi$pers, phi$net), find_phi(phi$pers, phi$search),
          'Dr',     NA, NA, find_phi(phi$dr, phi$forum), find_phi(phi$dr, phi$net), find_phi(phi$dr, phi$search),
          'Online Forum',  NA, NA, NA, find_phi(phi$forum, phi$net), find_phi(phi$forum, phi$search),
          'Social Networking Site',    NA, NA,NA, NA, find_phi(phi$net, phi$search),
          'Online Search', NA, NA, NA, NA, NA)

kable(phi_squared_tests, digits=3,  
      caption = "Phi Coefficient ($\\phi$) of binary association between each Information Search Vehicle",
      col.names = c("",'Person', 'Dr', 'Online Forum', 'Social Networking Site', 'Online Search')) %>% 
  kable_styling(latex_options = c("hold_position", "striped"),
                full_width = F)

```

To further investigate how people search for information, I asked those who had
sought information out intentionally, "How did you look for information about
the Covid-19 vaccine?" Because respondents could select as many responses as
applicable, each response is recorded as a dichotomous variable. I first
investigate how each Information Search Vehicle are related to each other using
the Phi Coefficient ($\\phi$) of binary association between each Information
Search Vehicle pair. The Phi coefficient can be interpreted like a correlation
coefficient, with numbers closer to -1 or positive one indicating a very strong
relationship [@warrens08]. Table \@ref(tab:model-2-chi2) shows the results of
these coefficients. Out of the five information search vehicles investigated,
there is evidence that the vehicles are distinctly different concepts because
most coefficients fall under 0.3, the general rule of thumb for correlation 
coefficients to indicate a relationship. There does seem to be a relationship
between Online Forums and Social Networking Sites, possibly due to an
individual's propensity to use online networks over in-person networks. These
coefficients indicate a weak relationship between any two vehicles of
information search and and that respondents in my sample utilize each of these a
little bit differently.

```{r table-model-2}
#| results = "asis"

part2_models <- list()

part2_models[['Doctor']] <- glm(formula = int_sought_fr_dr ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Person']] <- glm(formula = int_sought_fr_pers ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Online Forum']] <- glm(formula = int_sought_fr_onlgroup ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Social Networking Site']] <- glm(formula = int_sought_fr_onlnet ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

part2_models[['Online Search']] <- glm(formula = int_sought_fr_onlsearch ~ 
           receive_i_fr_dr + receive_i_fr_pers + receive_i_fr_tv +    
           receive_i_fr_onlnet + receive_i_fr_onlgroup + 
           age + gender + college +
           race_white + race_black + race_nat + race_asian + hispanic,  
         data = rain, family = binomial(link="logit"))

moda <- glm(formula = int_sought_fr_dr ~ 1,
            data = rain, family = binomial(link="logit"))
r2_a <- as.numeric(1-(logLik(part2_models[['Doctor']])/logLik(moda)))

modb <- glm(formula = int_sought_fr_pers ~ 1,
            data = rain, family = binomial(link="logit"))
r2_b <- as.numeric(1-(logLik(part2_models[['Person']])/logLik(modb)))

modc <- glm(formula = int_sought_fr_onlgroup ~ 1,
            data = rain, family = binomial(link="logit"))
r2_c <- as.numeric(1-(logLik(part2_models[['Online Forum']])/logLik(modc)))

modd <- glm(formula = int_sought_fr_onlnet ~ 1,
            data = rain, family = binomial(link="logit"))
r2_d <- as.numeric(1-(logLik(part2_models[['Social Networking Site']])/logLik(modd)))

mode <- glm(formula = int_sought_fr_onlsearch ~ 1,
            data = rain, family = binomial(link="logit"))
r2_e <- as.numeric(1-(logLik(part2_models[['Online Search']])/logLik(mode)))

rows <- tribble(~term, ~Doctor, ~Person, ~`Online Forum`, ~`Social Networking Site`, ~`Online Search`,
                "$R_{McFadden}^2$", r2_a,   r2_b, r2_c, r2_d, r2_e)
attr(rows, 'position') <- c(14, 9)

modelsummary(part2_models,
             align="llllll",
             statistic = NULL,
            estimate = "{estimate}{stars}",
            gof_omit = 'AIC',
            coef_map = regression_labels,
            notes = list('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Raw, Unexponentiated Coefficients'),            
            title = 'What predicts someone intentionally searched for information, by vehicle of information search',
            # exponentiate = TRUE,
            add_rows = rows,
            output = "kableExtra") %>%
  kable_styling(latex_options = c("scale_down", "hold_position", "striped"))  %>%
  add_header_above(c(" " = 1, "Dependent Variable" = 5))
```

Results of each logistic regression are displayed in Table \@ref(tab:table-model-2).
Modeling each response as a separate formula is useful for interpretation, but
there are important caveats: because each logistic regression is on a different
scale, one must not compare the magnitude of coefficients between models. We
can, however, use the direction and significance of the coefficients to draw
conclusions about these different forms of information search vehicles.

The first model in Table \@ref(tab:table-model-2) predicts whether or not a
respondent sought information intentionally about Covid-19 vaccinations by
asking their doctor or another health professional. First and foremost, I find
that those who received information about Covid-19 vaccinations were
`r round(exp(part2_models$Doctor$coefficients[['receive_i_fr_dr1']]), 2)` times
more likely to seek out more information from their doctor than those who had 
not ( _p_ < 0.001). However, receiving information from other vehicles also 
increased the odds of seeking out information from medical professions. Those who
received information from a person like a friend or a relative were 
`r round(exp(part2_models$Doctor$coefficients[['receive_i_fr_pers1']]), 2)` times
more likely to ask their doctor ( _p_ < 0.001), while those who received 
information from the news or television were
`r round(exp(part2_models$Doctor$coefficients[['receive_i_fr_tv1']]), 2)`
times more likely ( _p_ < 0.01). I also find that those who received information
from an online forum like a subreddit were 
`r label_percent()(round(exp(part2_models$Doctor$coefficients[['receive_i_fr_onlgroup1']]), 2)-1)`
more likely to ask their doctors than those who hadn't ( _p_ < 0.05). Finally, 
respondents with an associate's degree or higher were 
`r label_percent()(round(exp(part2_models$Doctor$coefficients[['college1']]), 2)-1)`
more likely to ask their doctor for information on the Covid-19 vaccinations
than those who held less education ( _p_ < 0.01).

The second model predicts whether or not a respondent sought information about
Covid-19 vaccinations by asking a person like friend, neighbor, or family member
that they know. Only three coefficients in this model were significant and help
us draw any conclusions on who is most likely to activate their personal social
network for informational support. In this study, I find that receiving
information about the vaccine from a doctor, their personal network, or a social
networking site is associated with seeking out more information. Those who 
received information from a doctor or other medical professional were 
`r round(exp(part2_models$Person$coefficients[['receive_i_fr_dr1']]), 2)`
times more likely to seek information from a person like a friend  ( _p_ < 0.001). 
Moreover, it seems that receiving information from your personal social network,
like a friend or relative, is associated with seeking information from those 
same people; in fact, I find that those who did are 
`r round(exp(part2_models$Person$coefficients[['receive_i_fr_pers1']]), 2)` more
likely to do so ( _p_ < 0.001). Finally, the respondents who received information
about the vaccine on a social networking site were 
`r label_percent()(round(exp(part2_models$Person$coefficients[['receive_i_fr_onlnet1']]), 2)-1)`
 more likely to seek information from their personal network than those who hadn't 
( _p_ < 0.01). The unexpected result of this model is that very few factors predict
personal social network activation and this model has the lowest McFadden's R^2, 
indicating possible omitted variables.

I also wanted to see what was associated with posting in an online discussion
group, listserve, or other online forum like a Facebook Group or Subreddit about
Covid-19 vaccinations. This is conceptually different that posting a query on a
social networking site like Facebook or twitter where identities are involved,
but is theoretically more [costly?][friction?] than searching on an internet
search engine. I found that having received information about the vaccine from 
your personal social network (like a friend or relative), the news, or an online
forum was associated with higher odds of querying an online forum. Specifically,  
receiving information from your personal network is associated with 
`r round(exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_pers1']]), 2)`
higher odds ( _p_ < 0.001), from news or television media with `r round(exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_tv1']]), 2)` higher odds ( _p_ < 0.05), and 
receiving information on an online forum associated with 
`r round(exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_onlgroup1']]), 2)`
higher odds ( _p_ < 0.001). Moreover, I found that those with an associate's 
degree or above were 
`r round(exp(part2_models$'Online Forum'$coefficients[['college1']]), 2)` times
more likely to search for information on an online forum ( _p_ < 0.001). Finally,
I found a racial-ethnic with Hispanic-Latinos; I found that they were about 
`r label_percent()(exp(part2_models$'Online Forum'$coefficients[['hispanic1']])-1)`
more likely to query an online forum than non-Hispanics ( _p_ < 0.05). However, 
there were also a few variables that lowered the odds of posting in an online 
discussion group. First, those who had received information from a social 
networking site were
`r label_percent()(1-exp(part2_models$'Online Forum'$coefficients[['receive_i_fr_onlnet1']]))` 
less likely to query a forum than those who hadn't (relative odds ratio =
$1 - exp(\beta k)$, _p_ < 0.01). Morever, I find more ethno-racial effects: 
Respondents who claimed Asian ancestry were 
`r label_percent()(1-exp(part2_models$'Online Forum'$coefficients[['race_asian1']]))` 
less likely to query online forum ( _p_ < 0.01).

The fourth model predicts whether an individual sought information about 
Covid-19 vaccinations through posted queries on a social networking site such 
as Facebook timeline, Twitter status update, or Linkedin. Receiving information
about Covid-19 seems to be quite related to posting on socia media. Those who 
received information from their doctor's had 
`r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_dr1']]), 2)`
higher odds of posting on a social networking site ( _p_ < 0.01), while those 
who received their information from tv/news had `r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_tv1']]), 2)` higher odds ( _p_ < 0.05). 
Furthermore, those who received information on social networking sites were more 
likely to post queries on social networking sites ($\beta$ = 
`r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_onlnet1']]), 2)`,  _p_ < 0.001). 
The respondents in our sample who received information on online forums had
`r round(exp(part2_models$'Social Networking Site'$coefficients[['receive_i_fr_onlgroup1']]), 2)` 
higher odds of posting queries on their social networking platforms ( _p_ < 0.001). 
I found a relationship between education and posting on social networking sites
as well; specifically, those with an Associate's degree or above were `r round(exp(part2_models$'Social Networking Site'$coefficients[['college1']]), 2)` times more likely to post vaccination 
queries on social networking sites than those with lower education ( _p_ < 0.001).
Finally, I find that Hispanic-Latinos were 
`r label_percent()(exp(part2_models$'Social Networking Site'$coefficients[['hispanic1']])-1)`
more likely to ask their connections on a social networking site about the 
Covid-19 vaccinations than non-Hispanics ( _p_ < 0.01).

Searching for information using an online search engine such as Google or Bing 
is theorically the search vehicle with the least amount of social or cognitive 
friction. Our descriptive statistics showed that not everyone used this vehicle,
with only 41% of our sample (See Table \@ref(tab:table-1-desc)) having sought
information about the Covid-19 vaccines using online search engines. My model
reveals interesting patterns to help explain the variation. As with previous 
models, receiving information of any kind increase the odds of online search, 
though to varying magnitudes (Doctor:
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_dr1']]), 2)` odds; 
Person:
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_pers1']]), 2)` odds; 
News or Television: 
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_tv1']]), 2)` odds; 
Social Networking Site:
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_onlnet1']]), 2)` odds; 
Online Forum: 
`r round(exp(part2_models$'Online Search'$coefficients[['receive_i_fr_onlgroup1']]), 2)` odds).
I also find that older respondents were more likely to search online, with each 
addition year aged to yield
`r label_percent()(exp(part2_models$'Online Search'$coefficients[['age']])-1)`
higher odds of searching online ( _p_ < 0.001). The most interesting variations
come from what decreases the odds of online search. First, those with an
Associate's degree or higher were 
`r label_percent()(1-exp(part2_models$'Online Search'$coefficients[['college1']]))` 
less likely to search online for questions about Covid-19 than those with less
educational attainment ( _p_ < 0.001). In addition, I find that Hispanic-Latinos 
had `r label_percent()(1-exp(part2_models$'Online Search'$coefficients[['hispanic1']]))` 
lower odds of searching online than non-Hispanics ( _p_ < 0.05).

## Vaccination Views

```{r table-model-3}
#| results = "asis"

mod3 <- glm(formula = vacc_view_pos ~ receive_i_fr_dr +   
           receive_i_fr_pers +      
           receive_i_fr_tv +    
           receive_i_fr_onlnet +    
           receive_i_fr_onlgroup + 
           int_sought_fr_pers +  
           int_sought_fr_dr +    
           int_sought_fr_onlgroup +
           int_sought_fr_onlnet +
           int_sought_fr_onlsearch +
           age + 
           gender + 
           race_white +   
           race_black +    
           race_nat +      
           race_asian + hispanic +     
           college,  
         data = rain, family = binomial(link="logit"))

nullmod3 <- glm(formula = vacc_view_pos ~ 1,
            data = rain, family = binomial(link="logit"))
psuedor2_mod3 <- as.numeric(1-(logLik(mod3)/logLik(nullmod3)))

rows <- tribble(~term, ~model1,
                "$R_{McFadden}^2$", psuedor2_mod3)
attr(rows, 'position') <- c(19, 4)

modelsummary(mod3,
            gof_omit = 'AIC', statistic = NULL,
            estimate = "{estimate}{stars} ({std.error})",
            title = 'Does receiving or searching for information predict vaccination status?',
            notes = list('* p < .05. ** p < .01. *** p < .001 (two-tailed test).',
                      'Raw, Unexponentiated Coefficients'),            
            coef_map = regression_labels,
            add_rows = rows,
            # exponentiate = TRUE,
            output = "kableExtra") %>%
  kable_styling(latex_options = c("hold_position", "striped"))
```
Finally, in Table \@ref(tab:table-model-3), I investigate whether there is a
relationship between information receiving and information seeking behaviors and
the choice to get vaccinated against Covid-19 through a logistic regression
model predicting Vaccination Outlook. I find that receiving information from a
doctor or other medical profession is associated with 
`r round(exp(mod3$coefficients[['receive_i_fr_dr1']]), 2)` odds of receiving 
a Covid-19 vaccinated ( _p_ < 0.001) and that seeking information from a doctor
also increases the odds of vaccination by 
`r round(exp(mod3$coefficients[['int_sought_fr_dr1']]), 2)` ( _p_ < 0.001).
I find that my respondents who received information on a social networking site 
like Facebook or Twitter about the vaccines were 
`r label_percent()(1-round(exp(mod3$coefficients[['receive_i_fr_onlnet1']]), 2))` 
**less likely** to get vaccinated than those who received no information through
that vehicle ( _p_ < 0.001). However, I did find that those who reached out to
family or friends with their concerns were 
`r label_percent()(exp(mod3$coefficients[['int_sought_fr_pers1']])-1)` more likely
to become vaccinated ( _p_ < 0.05). Furthermore, my respondents who held an 
Associate's degree or above had `r round(exp(mod3$coefficients[['college1']]), 2)`
higher odds of becoming immunized ( _p_ < 0.001). And finally, I found further
ethno-racial differences in my sample. I found that Hispanic-Latinos had 
`r round(exp(mod3$coefficients[['hispanic1']]), 2)` higher odds of vaccination
than non-Hispanics ( _p_ < 0.05) and that my Asian sample had 
`r round(exp(mod3$coefficients[['race_asian1']]), 2)` higher odds as well over
non-Asians ( _p_ = 0.09). See Figure \@ref(fig:plot-model-3) for a visual 
representation of how these different coefficients are related to vaccination 
outlook. 
<!-- TODO: watch out about saying vaccinated, this variable is about opinion -->

```{r plot-model-3}
#| fig.cap='Plot of Coefficients, Model 3'

plot_summs(mod3,
           scale = TRUE,
           coefs =  c(
             'Received Info, Doctor' = 'receive_i_fr_dr',
             'Received Info, Person' = 'receive_i_fr_pers',
             'Received Info, News' = 'receive_i_fr_tv',
             'Received Info, Social Networking Site' = 'receive_i_fr_onlnet',
             'Received Info, Online Forum' = 'receive_i_fr_onlgroup',
             'Sought Info' = 'int_sought',
             'Sought Info, Doctor' = 'int_sought_fr_dr',
             'Sought Info, Person' = 'int_sought_fr_pers',
             'Sought Info, Social Networking Site' = 'int_sought_fr_onlnet',
             'Sought Info, Online Forum' = 'int_sought_fr_onlgroup',
             'Sought Info, Online Search' = 'int_sought_fr_onlsearch',
             'Age' = 'age',
             'Gender, Male' = 'gender=Male',
             'Gender, Other' = 'gender=Other',
             "Associate's Deg or above" = 'college',
             'White' = 'race_white',
             'Black' = 'race_black',
             'Native American' = 'race_nat',
             'Asian' = 'race_asian',
             'Hispanic or Latino/x' = 'hispanic'),
           robust = TRUE) +
  labs(title = 'Does receiving or searching for information predict vaccination sentiment?', 
       caption = 'Exponentiated Coefficients: Scaled and Robust') + 
  ylab("") +
  theme_diss()+
  scale_color_manual(values = my_palette[3])
```

# Discussion
Previous research in Sociology and Infodemiology has greatly failed to
distinguish between the activation of information seeking behaviors online and
offline. In this paper, I aim to uncover just how different various methods of
information search are and begin a line of research inquiry that investigates
the factors associated with each different vehicle. Finally, I also aimed to
uncover how the different fountains of information and information search
vehicles hold real world consequences through their associations with Covid-19
vaccination rates and intentions.

I first find that receiving information about Covid-19 vaccinations from any
outlined source increases the likelihood of searching out more information. I
also find important demographic differences in the overall propensity to seek
out information. First, I find that the older a respondent is, the more likely
they are to seek out information about Covid-19 vaccinations, possibly
indicating [???? TODO]. Furthermore, I find that those who are more educated,
that is my respondents who hold an Associate's degree or higher were more likely
to seek out information; for the most part this trend can be observed in Table
\@ref(tab:table-model-2) as well. Human Capital Theory [@mirowsky_ross98] may
help to explain this finding; this theory claims that education helps
individuals gaining health-related knowledge and use this knowledge to be
proactive in their own lives. Finally, I find that my respondents who identified
as Black were more likely to seek out further information; this may be explained
by the maltreatment that African-Americans have endured in the US
healthcare system [@baileyStructuralRacismHealth2017] which has led to major
distrust of the medical establishment among the population [@center2019a; @murray15].

(May need to cite [@bronson2014don] above or elsewhere)

table 2 main findings

usage of different information search vehicles varies by demographics and also reception of different ifnromation
- overall, receiving information about Covid-19 vaccination generally spurs information seeking behaviors, though there are variations
- receive info from doctor leads to searching info from doctor, likely indicates good communication and trust in doctor, solid relationship with healthcare professionals
- few things predict seeking out information from another person, personal networks are so embedded in everyday life so perhaps don't remember doing it or behavioral prediction is notoriously hard
- Received Info, Social Networking Site negatively predicts Online Forum, indicating that if your social networks fulfill your need for information you don't need to extend out to online forums to ask for things, need is already met? 
- Asians are really unlikely to use online forums to find information in this context, very strong association but hispanics are more likely than non-hispanics to search online 
- online search is much less likely from those with higher education, which is counter intuitive since they are the population that has been the most trained in research methods and critical thinking .. perhaps they trust the information that they hear if they consider it to be a trusted source (typical higher education format of trust the professor) so they don't look deeper into it? basically already have high social capital in other ways, already set up to receive the information they needed and trusted
- older populations are more likely to 'just google it'

takeaways: Not everyone is activating their personal social networks or using Google in the same rates, the internal variation of users is important to consider when using something like Google Trends
education really affects information seeking behavior
ethno-racial differences affect information seeking behavior, more research should look into why there are group differences (likely cultural)

Takeaway 2: in person, online, or search engines?

table 3 main findings 
- very important for getting people to get vaccinated - a major goal for public Health authorities over the last two years
- extremely importnat given misinformation around vaccinations 
- receiving information from a doctor or other medical profession (or seeking) leads to more positive opinions. action inspired by this - difficult because of lack of primary care doctors for many americans, but every doctor should discuss the importance for every patient they interact with 
- receiving information from a social networking site NOT an online for results in less vaccinations. 
- differences in education level and race ; seen in prior literature however most current literature claims that anti-vacination attitudes span social class, political ideology, and X.

social netwokring site
info from doctor

<!-- this never before seen pattern in sociology may be a result of response bias, (most likely to select more information) but it also could indicate a sheltering effect ... <--- I wish I knew what I was talking about here -->

rather than on “the receiver as an active information seeker and processor."
[@johnsonComprehensiveModelCancerRelated1993].

Est velit egestas dui id ornare arcu odio ut. Tempus iaculis urna id volutpat
lacus laoreet non curabitur gravida. Integer malesuada nunc vel risus commodo.
Ipsum nunc aliquet bibendum enim facilisis gravida. Ac turpis egestas maecenas
pharetra convallis posuere morbi leo. Id diam vel quam elementum pulvinar etiam.

# Conclusion
Previous research has greatly failed to distinguish between the activation of
information seeking behaviors online and offline. In this paper, I aim to
uncover just how different various methods of information search are and begin a
line of research inquiry that investigates the factors associated with each vehicle. 
Finally, I also aimed to uncover how the different fountains of information and
information search vehicles hold real world consequences through their associations
with Covid-19 vaccination rates and intentions.

The vaccines against the COVID-19 coronavirus prevent severe illness,
hospitalization, and death. However, as I’m writing this, less than 55% of
Arizonans over 12 years old are fully vaccinated against the virus. We know that
people exposed to misinformation are much less likely to receive vaccinations
against the virus [@loombaMeasuringImpactCOVID192021]. It is not an exaggeration
to claim that misinformation has been deadly for Arizona
[@pathakInfodemicsCOVID19Role2020].

In academia and policy, the main focus is on the information that is sent to
consumers, seeing people as passive receivers of information. My dissertation
instead focuses on individuals as active agents in their search for information
which helps me uncover the information individuals are exposed to. By looking at
how individuals search, I aim to identify pain points and focus areas for future
interventions in the misinformation process.

This research project is important because the vehicles used to search by people
impacts the information they are exposed to. Given the large swaths of both
misinformation and disinformation regarding Covid-19 [@mottaHowRightLeaningMedia2020; @pathakInfodemicsCOVID19Role2020; @shahsavariConspiracyTimeCorona2020] and
the measurable impacts this misinformation has had on pandemic-related health
behaviors, how we choose to search for information affects our own lives and the
lives of others. Moreover, individuals who do not utilize social networks for
information search may feel like they have no network support, leading to
feelings of network isolation with major implications for mental health 
[@thoitsMechanismsLinkingSocial2011]

# References {-}

<div id="refs"></div>

\newpage
# Appendix 1: Survey Questions

-------------------------------------------------

*Q181*: Now, we're going to transition a bit to ask you about the information you received during the Covid-19 pandemic.

-------------------------------------------------

*Q170*: In the past 12 months, without searching for it, did you receive information about the Covid-19 vaccine from … (check all sources you received information from)

*	your doctor or other health professional?   (1) 
*	a person like friend, neighbor, or family member that you know ?  (9) 
*	From television news channel or the newspaper?  (10) 
*	From an online discussion group, listserve, or other online forum including a Facebook group or subreddit?  (11) 
*	From a social networking site such as Facebook timeline, Twitter status update, or linkedin?  (12) 
*	Other  (13) _______

-------------------------------------------------

*Q171*: And how about you yourself intentionally looking for information about a Covid-19 vaccine? Such information can include things such as advice, clarification, facts, and experiences.

*	Yes  (4) 
*	No  (5)

-------------------------------------------------

*Q172*: How did you look for information about the Covid-19 vaccine?

*	Asking a person like friend, neighbor, or family member that I know  (1) 
*	Asking my doctor or another health professional  (4) 
*	Posted queries in an online discussion group, listserve, or other online forum like a Facebook Group or Subreddit  (5) 
*	Posted queries on a social networking site such as Facebook timeline, Twitter status update, or Linkedin  (6) 
*	Searched for my question using an online search engine such as Google or Bing  (7) 
*	Other  (8) ___________

-------------------------------------------------

*Q173*: What sort of information did you search for? 
 Separate different topics with a comma or semi-colon.

-------------------------------------------------

*Q174*: How useful was the information you found?

*	Extremely useful  (22) 
*	Very useful  (23) 
*	Moderately useful  (24) 
*	Slightly useful  (25) 
*	Not at all useful  (26)

-------------------------------------------------

*Q175*: Did the information you learned affect your decision to get vaccinated against Covid-19?

*	Yes  (39) 
*	No  (40)

-------------------------------------------------

*Q179*: Did you receive a Covid-19 vaccine?

*	Yes  (9) 
*	No  (10) 
*	I'm unsure or would not like to respond  (11)

-------------------------------------------------

*Q180*: Do you plan to receive a vaccine for the prevention of the Covid-19 virus?

*	Definitely not  (9) 
*	Probably not  (10) 
*	Might or might not  (11) 
*	Probably yes  (12) 
*	Definitely yes  (13) 
*	I would not like to respond  (14)

-------------------------------------------------

*gender*: What is your gender?

*	Male  (1) 
*	Female  (2) 
*	Other  (3) ________________
*	Prefer not to say  (4)

-------------------------------------------------

*hispanic*: Are you Hispanic, Latino/a/x, or Latin American Origin?

*	Yes  (1) 
*	No  (2)

-------------------------------------------------

*race*: What is your race? If you are "mixed race", select all that apply.

*	White  (1) 
*	Black or African American  (2) 
*	American Indian or Alaskan Native  (3) 
*	Asian (please specify):  (4) _______
*	Other (please specify):  (5) _______
*	Prefer not to say  (6)

-------------------------------------------------

*educ*: What is the highest level of education that you have completed?

*	Less than high school  (1) 
*	High school graduate  (2) 
*	Some college  (3) 
*	Associate's or Technical degree  (4) 
*	Bachelor's degree  (5) 
*	Graduate or professional degree  (6)

-------------------------------------------------

*income*: We would be interested to know roughly what your total household income
before taxes is. We mean income from all sources, including welfare, stock
dividends, other household members' income, etc. In 2020, in which bracket did your total family income fall?

*	Under $1,000  (1) 
*	\$1,000 to \$9,999   (2) 
*	\$10,000 to \$19,999  (3) 
*	\$20,000 to \$29,999  (4) 
*	\$30,000 to \$39,999  (5) 
*	\$40,000 to \$49,999  (6) 
*	\$50,000 to \$59,999  (7) 
*	\$60,000 to \$74,999  (8) 
*	\$75,000 to \$89,999   (9) 
*	\$90,000 to \$109,999  (10) 
*	\$110,000 to \$129,999  (11) 
*	\$130,000 to \$149,999  (12) 
*	\$150,000 or over  (13) 
*	Don't know  (14)

-------------------------------------------------

*age*: What is your age? (in years)